### 2016-01-13 paper/ms.tex
start idea dump of text

\documentclass[]{openjournal}

\bibliographystyle{apj}
\usepackage{graphicx}
\usepackage[suffix=]{epstopdf}
\usepackage[export]{adjustbox}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{url}
\usepackage{xspace}

%    Make Scientific Notation
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

% make the word Kepler italicized
\newcommand{\Kepler}{\textsl{Kepler}\xspace}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%
\title{}

\shorttitle{}
\shortauthors{Davenport et al.}

\author{
	James R. A. Davenport\altaffilmark{1,2,3}
	}

\altaffiltext{1}{Corresponding author: James.Davenport@wwu.edu}
\altaffiltext{2}{Department of Physics \& Astronomy, Western Washington University, Bellingham, WA 98225}
\altaffiltext{3}{NSF Astronomy and Astrophysics Postdoctoral Fellow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
The idea is simple: ET's would place artificial satellites in orbit around stars surrounding their home star system. The orbit period of the remote artificial satellites (or beacons) would be proportional to the distance from the remote star to the home star system. Each remote beacon system would be part of a network map leading to the home system. If the orbital period vs beacon distance relationship was known, the exact location of the home system could be triangulated using only a subset of the beacons.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Previous work has suggested looking for SETI signals from interstellar ``beacons''
\cite{benford2008}
However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}. This is every expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}

Instead, much cheaper to block light, rather than shine it.
\cite{arnold2005}
on the feasibility of transits for use in SETI detection, and 
\cite{arnold2005a}
on the impact of artificial structures on transits. also this work on that:
\cite{wright2016}.
recent data from \Kepler \citep{borucki2010}, has found weird transit-like signals
\cite{boyajian2015}, 
which some have considered as SETI candidates. However, follow-up observations have yet to yield any confirming signatures, and instead this looks like a YSO w/ comets maybe \citep{lisse2015}.

\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin. However, very little information can be effectively transmitted using eclipses, even with extreme precision in our recovery.
In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization. This system has the advantage of being potentially detectable in the near future via exoplanet searches.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinated Transits as ET Beacons}

\begin{figure}[]
\centering
\includegraphics[width=3.5in]{../figures/dist_per.png}
\caption{schematic figure of the signal to detect in 1 dimension}
\label{fig:1d}
\end{figure}

\begin{figure}[]
\centering
\includegraphics[width=3.5in]{../figures/sky_per.png}
\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
\label{fig:2d}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Predictions for Upcoming Surveys}
TESS would be great for this in terms of spatial-temporal coverage.

LSST very good for finding SETI signals of this geometric style, but not ideal for events requiring such dedicating monitoring

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}

%%%%%%%%%%%%%%%%%
\acknowledgments
JRAD is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST-1501418.

\bibliography{/Users/james/research/references}

\end{document}

### 2016-01-22 paper/ms.tex
abstract drafted

@@
-\documentclass[]{openjournal}
+%\documentclass[]{openjournal}
+\documentclass[iop]{emulateapj}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{}
+\title{SETI in the Spatial-Temporal Domain: Spatially Coordinated Transits as an ET Signal}
 
-\shorttitle{}
+\shorttitle{SETI in the Spatial-Temporal Domain}
 \shortauthors{Davenport et al.}
@@
 \begin{abstract}
-The idea is simple: ET's would place artificial satellites in orbit around stars surrounding their home star system.
-The orbit period of the remote artificial satellites (or beacons) would be proportional to the distance from the remote star to the home star system.
-Each remote beacon system would be part of a network map leading to the home system.
-If the orbital period vs beacon distance relationship was known, the exact location of the home system could be triangulated using only a subset of the beacons.
+Traditional searches for extraterrestrial intelligence (SETI) focuses on temporal or multi-wavelength studies of single stars to detect transient or excess photon emission from artificial sources.
+However, the latest generation of synoptic time domain surveys enables a spatial-temporal SETI, where signals originate from resolved sources or multiple stars.
+Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
+Artificial planets would act as beacons, being placed in orbit around stars surrounding a home star system.
+The orbital period of each artificial satellite would be proportional to the distance from the beacon star to the home star system.
+If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transit beacons.
+Current and future exoplanet surveys may be able to detect such spatially coordinated transits around nearby low-mass stars.
+Importantly, the spatial-temporal domain deserves increased attention from SETI researchers.
 \end{abstract}
@@
 
+In the search for extraterrestrial intelligence (SETI), most approaches focus on direct detection of photons originating from transmission or waste energy.
+These searches occur over a range of wavelengths, and require extraterrestrials to produce sufficient radiation to be detected apart from their parent star.
+While this may be feasible at radio wavelengths, it is much more difficult in the optical regime where much of our time domain surveys operate.
+However,  it is more simple to block significant light from a star than to produce enough to out shine it, which has led to recent studies of time domain data for signatures of transiting artificial structures.
+
 Previous work has suggested looking for SETI signals from interstellar ``beacons''
@@
 However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
-This is every expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
+This is a very expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
 
@@
 \cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
-However, very little information can be effectively transmitted using eclipses, even with extreme precision in our recovery.
+However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
+
 In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization.

### 2016-09-19 paper/ms.tex
change paper style

@@
 %\documentclass[]{openjournal}
-\documentclass[iop]{emulateapj}
+\documentclass[manuscript, letterpaper]{aastex6}
+\bibliographystyle{aasjournal}
 
@@
 \usepackage[suffix=]{epstopdf}
-\usepackage[export]{adjustbox}
 \usepackage{natbib}
@@
 \usepackage{xspace}
+
+% from here: https://github.com/dfm/peerless/blob/master/document/ms.tex#L19-L69
+% ----------------------------------- %
+% start of AASTeX mods by DWH and DFM %
+% ----------------------------------- %
+
+\setlength{\voffset}{0in}
+\setlength{\hoffset}{0in}
+\setlength{\textwidth}{6in}
+\setlength{\textheight}{9in}
+\setlength{\headheight}{0ex}
+\setlength{\headsep}{\baselinestretch\baselineskip} % this is 2 lines in ``manuscript''
+\setlength{\footnotesep}{0in}
+\setlength{\topmargin}{-\headsep}
+\setlength{\oddsidemargin}{0.25in}
+\setlength{\evensidemargin}{0.25in}
+
+\linespread{0.54} % close to 10/13 spacing in ``manuscript''
+\setlength{\parindent}{0.54\baselineskip}
+\hypersetup{colorlinks = false}
+\makeatletter % you know you are living your life wrong when you need to do this
+\long\def\frontmatter@title@above{
+\vspace*{-\headsep}\vspace*{\headheight}
+\noindent\footnotesize
+{\noindent\footnotesize\textsc{\@journalinfo}}\par
+{\noindent\scriptsize Preprint typeset using \LaTeX\ style AASTeX6\\
+With modifications by David W. Hogg and Daniel Foreman-Mackey
+}\par\vspace*{-\baselineskip}\vspace*{0.625in}
+}%
+\makeatother
+
+% Section spacing:
+\makeatletter
+\let\origsection\section
+\renewcommand\section{\@ifstar{\starsection}{\nostarsection}}
+\newcommand\nostarsection[1]{\sectionprelude\origsection{#1}}
+\newcommand\starsection[1]{\sectionprelude\origsection*{#1}}
+\newcommand\sectionprelude{\vspace{1em}}
+\let\origsubsection\subsection
+\renewcommand\subsection{\@ifstar{\starsubsection}{\nostarsubsection}}
+\newcommand\nostarsubsection[1]{\subsectionprelude\origsubsection{#1}}
+\newcommand\starsubsection[1]{\subsectionprelude\origsubsection*{#1}}
+\newcommand\subsectionprelude{\vspace{1em}}
+\makeatother
+
+\widowpenalty=10000
+\clubpenalty=10000
+
+\sloppy\sloppypar
+
+% ------------------ %
+% end of AASTeX mods %
+% ------------------ %
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{SETI in the Spatial-Temporal Domain: Spatially Coordinated Transits as an ET Signal}
+\title{SETI in the Spatial-Temporal Domain}
 
 \shorttitle{SETI in the Spatial-Temporal Domain}
-\shortauthors{Davenport et al.}
+\shortauthors{Davenport}
 
 \author{
-	James R. A. Davenport\altaffilmark{1,2,3}
+	James R. A. Davenport\altaffilmark{1,2}
 	}
 
-\altaffiltext{1}{Corresponding author: James.Davenport@wwu.edu}
-\altaffiltext{2}{Department of Physics \& Astronomy, Western Washington University, Bellingham, WA 98225}
-\altaffiltext{3}{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+%\altaffiltext{1}{Corresponding author: James.Davenport@wwu.edu}
+\altaffiltext{1}{Department of Physics \& Astronomy, Western Washington University, Bellingham, WA 98225}
+\altaffiltext{2}{NSF Astronomy and Astrophysics Postdoctoral Fellow}
 
@@
 
+optical seti
+http://arxiv.org/abs/astro-ph/0506758
+
 Instead, much cheaper to block light, rather than shine it.
@@
 
+active and passive SETI in coordination (both temporal and spatial) with galactic events like supernovae 
+\citep{lemarchand1994}
+
+lighthouses
+\citep{zuluaga2015}
+
 \cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Coordinated Transits as ET Beacons}
+\section{Example: Coordinated Transits as ET Beacons}
+imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
+
+give the detailed example i have thought up.
+
+becons placed at distances
 
@@
 
+for clarity, this is what the ideal signal might look like on the sky
+
 \begin{figure}[]
@@
 
+%and then what it might look like in 3d, projected in to 2d
+
+%%%
+
+this type of beacon network is advantageous because it points directly back towards their home. only a few systems actually need to be transiting from any given line of sight.
+
+NOW THE FULL SIMULATION...
+- computation to do: if had 100 beacons, each placed at random orbital alignment, in 3d sphere around home system.
+- assume G stars, 
+- odds of observing transit of a fixed sized object versus orbital distance... goes down. need that plot to figure out probability. 
+- assume ET places beacons with uniform RADIAL density in 3d space out to some maximum distance (even \# of systems as function of radial distance) in bins of 10pc out to 100 pc (i.e.
+10 beacons in each 10pc bin)
+- orbital period is exact for each system, no bins of period
+- do Monte Carlo sim with these parameters to figure out how many transiting systems we'd observe
+
+---- make the plot for one MC realization of RA,Dec.... open circles for systems with no observed transits, colored for transits
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Predictions for Upcoming Surveys}
+\section{Prospects for Future Surveys}
+
+to get accurate census you need a complete time domain survey at some transit depth sensitivity out to some orbital period. going off our simulation, TESS might work for very short period planets
+
 TESS would be great for this in terms of spatial-temporal coverage.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary}
+\section{Discussion}
+this kind of distributed network could be very efficient at broadcasting the presence of a civilization. dust clouds could create sufficient "transit" signals, while not affecting orbital dynamics
 

### 2017-10-31 paper/ms.tex
change to modern aastex style

@@
-%\documentclass[]{openjournal}
-\documentclass[manuscript, letterpaper]{aastex6}
+\documentclass[modern]{aastex61}
 \bibliographystyle{aasjournal}
 
-\bibliographystyle{apj}
+%\bibliographystyle{apj}
 \usepackage{graphicx}
@@
 \usepackage{amsmath}
-\usepackage{url}
 \usepackage{xspace}
-
-% from here: https://github.com/dfm/peerless/blob/master/document/ms.tex#L19-L69
-% ----------------------------------- %
-% start of AASTeX mods by DWH and DFM %
-% ----------------------------------- %
-
-\setlength{\voffset}{0in}
-\setlength{\hoffset}{0in}
-\setlength{\textwidth}{6in}
-\setlength{\textheight}{9in}
-\setlength{\headheight}{0ex}
-\setlength{\headsep}{\baselinestretch\baselineskip} % this is 2 lines in ``manuscript''
-\setlength{\footnotesep}{0in}
-\setlength{\topmargin}{-\headsep}
-\setlength{\oddsidemargin}{0.25in}
-\setlength{\evensidemargin}{0.25in}
-
-\linespread{0.54} % close to 10/13 spacing in ``manuscript''
-\setlength{\parindent}{0.54\baselineskip}
-\hypersetup{colorlinks = false}
-\makeatletter % you know you are living your life wrong when you need to do this
-\long\def\frontmatter@title@above{
-\vspace*{-\headsep}\vspace*{\headheight}
-\noindent\footnotesize
-{\noindent\footnotesize\textsc{\@journalinfo}}\par
-{\noindent\scriptsize Preprint typeset using \LaTeX\ style AASTeX6\\
-With modifications by David W. Hogg and Daniel Foreman-Mackey
-}\par\vspace*{-\baselineskip}\vspace*{0.625in}
-}%
-\makeatother
-
-% Section spacing:
-\makeatletter
-\let\origsection\section
-\renewcommand\section{\@ifstar{\starsection}{\nostarsection}}
-\newcommand\nostarsection[1]{\sectionprelude\origsection{#1}}
-\newcommand\starsection[1]{\sectionprelude\origsection*{#1}}
-\newcommand\sectionprelude{\vspace{1em}}
-\let\origsubsection\subsection
-\renewcommand\subsection{\@ifstar{\starsubsection}{\nostarsubsection}}
-\newcommand\nostarsubsection[1]{\subsectionprelude\origsubsection{#1}}
-\newcommand\starsubsection[1]{\subsectionprelude\origsubsection*{#1}}
-\newcommand\subsectionprelude{\vspace{1em}}
-\makeatother
-
-\widowpenalty=10000
-\clubpenalty=10000
-
-\sloppy\sloppypar
-
-% ------------------ %
-% end of AASTeX mods %
-% ------------------ %
 
@@
 
-\author{
-	James R. A. Davenport\altaffilmark{1,2}
-	}
-
-%\altaffiltext{1}{Corresponding author: James.Davenport@wwu.edu}
-\altaffiltext{1}{Department of Physics \& Astronomy, Western Washington University, Bellingham, WA 98225}
-\altaffiltext{2}{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+\author{James. R. A. Davenport}
+\affiliation{Department of Physics \& Astronomy, Western Washington University, 516 High St., Bellingham, WA 98225, USA}
+\altaffiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+\affiliation{Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
+\altaffiliation{DIRAC Fellow}
 
@@
 \begin{abstract}
-Traditional searches for extraterrestrial intelligence (SETI) focuses on temporal or multi-wavelength studies of single stars to detect transient or excess photon emission from artificial sources.
-However, the latest generation of synoptic time domain surveys enables a spatial-temporal SETI, where signals originate from resolved sources or multiple stars.
+Traditional searches for extraterrestrial intelligence (SETI) focus on single stars to detect transient or excess photon emission from intelligent sources.
+However, the latest generation of synoptic time domain surveys enable a spatial-temporal SETI, where signals originate from spatially resolved sources or multiple stars.
 Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
-Artificial planets would act as beacons, being placed in orbit around stars surrounding a home star system.
+Artificial occulters would act as beacons, being placed in orbit around stars surrounding the central home star system.
 The orbital period of each artificial satellite would be proportional to the distance from the beacon star to the home star system.
@@
 Current and future exoplanet surveys may be able to detect such spatially coordinated transits around nearby low-mass stars.
-Importantly, the spatial-temporal domain deserves increased attention from SETI researchers.
+This example technique highlights the potential of the spatial-temporal domain for SETI researchers.
 \end{abstract}
+
+%\keywords{stars: activity --- stars: low-mass --- planets and satellites: detection}
 
@@
 \citep{lemarchand1994}
+coordinated times/places, also known as Schelling or Focal points \citep{schelling1960}, are the key to finding people in unknown time/space.
 
@@
 
+planets in Earth's transit zone \cite{heller2016}
+
 \cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
 However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
+
+\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
+
+a review of Schelling points from \citet{wright2017}
 
@@
 
-\begin{figure}[]
+\begin{figure*}[]
 \centering
-\includegraphics[width=3.5in]{../figures/dist_per.png}
-\caption{schematic figure of the signal to detect in 1 dimension}
-\label{fig:1d}
-\end{figure}
+\includegraphics[width=2.75in]{../figures/dist_per.pdf}
+\includegraphics[width=2.75in]{../figures/sky_per.pdf}
+\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
+\label{fig:2d}
+\end{figure*}
 
 for clarity, this is what the ideal signal might look like on the sky
-
-\begin{figure}[]
-\centering
-\includegraphics[width=3.5in]{../figures/sky_per.png}
-\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
-\label{fig:2d}
-\end{figure}
-
-%and then what it might look like in 3d, projected in to 2d
 
@@
 
+\begin{figure}[]
+\centering
+\includegraphics[width=3in]{../figures/3d_model.pdf}
+\caption{the model in 3 dimensions projected into the sky plane.
+1000 simulated systems that span the galactic distance of 5-50 pc (blue) with 10 recovered transits highlighted (red).
+these would be hot Jupiters at the short P end.}
+\label{fig:3d}
+\end{figure}
+
+\begin{figure}[]
+\centering
+\includegraphics[width=3in]{../figures/recov_fraction.pdf}
+\caption{fraction of hot Jupiter-like systems recovered in 1000 realizations of our 1000-star simulation.
+This curve is dependent on the ratio of the occulter-to-star size, and the orbital period range sampled.
+The average from our model is $10\pm 3$ systems recovered for these parameters.}
+\label{fig:recov}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+IDEA:
+could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
+How much more efficient would it be if we forced $i<10^\circ$, instead of $i<90^\circ$?
+
+for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
+by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
+
 %%%%%%%%%%%%%%%%%

### 2018-04-17 paper/ms.tex
a few tweaks on a plane ride

@@
-\documentclass[modern]{aastex61}
+\documentclass[modern]{aastex62}
 \bibliographystyle{aasjournal}
@@
 \begin{abstract}
-Traditional searches for extraterrestrial intelligence (SETI) focus on single stars to detect transient or excess photon emission from intelligent sources.
+Traditional searches for extraterrestrial intelligence (SETI) focus on single stars to detect excess or transient emission from intelligent sources.
 However, the latest generation of synoptic time domain surveys enable a spatial-temporal SETI, where signals originate from spatially resolved sources or multiple stars.
 Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
-Artificial occulters would act as beacons, being placed in orbit around stars surrounding the central home star system.
+In this scenario, artificial occulting objects would act as beacons, being placed in orbit around stars surrounding the central home star system.
 The orbital period of each artificial satellite would be proportional to the distance from the beacon star to the home star system.
-If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transit beacons.
-Current and future exoplanet surveys may be able to detect such spatially coordinated transits around nearby low-mass stars.
-This example technique highlights the potential of the spatial-temporal domain for SETI researchers.
+If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transiting beacons seen by an observer.
+Current and future exoplanet surveys would be able to detect such spatially coordinated transits around nearby low-mass stars.
+While contrived, this example highlights the need for SETI researchers to consider the spatial-temporal domain.
 \end{abstract}
@@
 In the search for extraterrestrial intelligence (SETI), most approaches focus on direct detection of photons originating from transmission or waste energy.
-These searches occur over a range of wavelengths, and require extraterrestrials to produce sufficient radiation to be detected apart from their parent star.
-While this may be feasible at radio wavelengths, it is much more difficult in the optical regime where much of our time domain surveys operate.
-However,  it is more simple to block significant light from a star than to produce enough to out shine it, which has led to recent studies of time domain data for signatures of transiting artificial structures.
+These searches occur over a range of wavelength regimes, and assume an extraterrestrial civilization will produce sufficient radiation to be detected apart from their parent star.
+While this may be feasible at radio wavelengths where Sun-like stars have lower emission, it is much more difficult to outshine the star in the optical regime where many present-day time domain surveys operate.
 
-Previous work has suggested looking for SETI signals from interstellar ``beacons''
-\cite{benford2008}
-However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
-This is a very expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
-
-optical seti
-http://arxiv.org/abs/astro-ph/0506758
-
-Instead, much cheaper to block light, rather than shine it.
-\cite{arnold2005}
-on the feasibility of transits for use in SETI detection, and 
-\cite{arnold2005a}
+Blocking a significant fraction of light from a star may instead be much easier than outshining it \cite{arnold2005}.
+This has led to recent searches for transit-like signatures that may have artificial origins.
+For example, PAPER demonstrated the observable impact of transits by non-spherical objects \cite{arnold2005a}.
 on the impact of artificial structures on transits. also this work on that:
@@
 
+Previous work has suggested looking for SETI signals from interstellar ``beacons''
+\cite{benford2008}
+However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
+This is a very expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
+\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
+However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
+\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
+
+%optical seti
+%http://arxiv.org/abs/astro-ph/0506758
+
+a review of Schelling points from \citet{wright2017}
+lighthouses \citep{zuluaga2015}
 active and passive SETI in coordination (both temporal and spatial) with galactic events like supernovae 
@@
 coordinated times/places, also known as Schelling or Focal points \citep{schelling1960}, are the key to finding people in unknown time/space.
+planets in Earth's transit zone \cite{heller2016} - maybe transits would be aligned for our viewing benefit?
 
-lighthouses
-\citep{zuluaga2015}
-
-planets in Earth's transit zone \cite{heller2016}
-
-\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
-However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
-
-\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
-
-a review of Schelling points from \citet{wright2017}
+{\bf THE POINT: } in this new spatial-temporal domain, we should be looking for things that happen in space, such as over densities of events, or coordination of events like transits.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Searching for Correlated Transits in Kepler}
+we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
+
+we find there is no correlation. Thats OK, but at least we searched
+
+\begin{figure}[]
+\centering
+\includegraphics[width=3in]{../figures/delta_per.png}
+\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
+median period differences are computed for bins of separation distance (orange points).
+there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
+\label{fig:2dcor}
+\end{figure}
+
+\begin{figure}[]
+\centering
+\includegraphics[width=3in]{../figures/delta_per2.png}
+\caption{correlation between transit periods and projected separations, as in Figure \ref{fig:2dcor}, but emphasizing the smallest separations in both axes.
+One outlier was identified in the similarity of Kepler 1292b and Kepler 1235b (black circle).
+However, these two systems are more than 600 pc separated in distance.}
+\label{fig:2dcor2}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Prospects for Future Surveys}
@@
 
+This research has made use of the Exoplanet Orbit Database and the Exoplanet Data Explorer at exoplanets.org.
+
 \bibliography{/Users/james/research/references}

### 2018-11-28 paper/ms.tex
update paper from yesterday

--- 2017-10-31 paper/ms.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{SETI in the Spatial-Temporal Domain}
+\title{SETI in the Optical Spatial-Temporal Domain}
 
-\shorttitle{SETI in the Spatial-Temporal Domain}
+\shorttitle{Spatial-Temporal SETI}
 \shortauthors{Davenport}
@@
 \author{James. R. A. Davenport}
-\affiliation{Department of Physics \& Astronomy, Western Washington University, 516 High St., Bellingham, WA 98225, USA}
-\altaffiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+%\affiliation{Department of Physics \& Astronomy, Western Washington University, 516 High St., Bellingham, WA 98225, USA}
+%\altaffiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
 \affiliation{Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
@@
 \begin{abstract}
-Traditional searches for extraterrestrial intelligence (SETI) focus on single stars to detect transient or excess photon emission from intelligent sources.
-However, the latest generation of synoptic time domain surveys enable a spatial-temporal SETI, where signals originate from spatially resolved sources or multiple stars.
-Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
-Artificial occulters would act as beacons, being placed in orbit around stars surrounding the central home star system.
+Traditional searches for extraterrestrial intelligence (SETI) or ``technosignatures'' focus on dedicated observations of single stars or regions in the sky to detect excess or transient emission from intelligent sources.
+The latest generation of synoptic time domain surveys  enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
+SETI methodology has not yet fully utilized such surveys, however, and the literature is in need of new search strategies.
+Here I describe the potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some new example SETI approaches.
+\end{abstract}
+
+%Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
+In this scenario, artificial occulting objects would act as beacons, being placed in orbit around stars surrounding the central home star system.
 The orbital period of each artificial satellite would be proportional to the distance from the beacon star to the home star system.
-If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transit beacons.
-Current and future exoplanet surveys may be able to detect such spatially coordinated transits around nearby low-mass stars.
-This example technique highlights the potential of the spatial-temporal domain for SETI researchers.
-\end{abstract}
+If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transiting beacons seen by an observer.
+Current and future exoplanet surveys would be able to detect such spatially coordinated transits around nearby low-mass stars.
+While contrived, this example highlights the need for SETI researchers to consider the spatial-temporal domain.
 
@@
 
-In the search for extraterrestrial intelligence (SETI), most approaches focus on direct detection of photons originating from transmission or waste energy.
-These searches occur over a range of wavelengths, and require extraterrestrials to produce sufficient radiation to be detected apart from their parent star.
-While this may be feasible at radio wavelengths, it is much more difficult in the optical regime where much of our time domain surveys operate.
-However,  it is more simple to block significant light from a star than to produce enough to out shine it, which has led to recent studies of time domain data for signatures of transiting artificial structures.
+Despite more than half a century of activity, the Search for Extra-Terrestrial Intelligence (SETI) is a field of study very much in its infancy. 
+\citet{wright2018c} note the volume of parameter space for SETI (or the search for a needle amongst the ``cosmic haystack'') at radio wavelengths has been barely explored, and that many obvious signals may indeed be awaiting our discovery.
+The lack of SETI activity by the astronomical community can be partially understood as due to funding limitations in recent decades, as well as social pressures against studying SETI experienced by professional astronomers and organizations \citep{wright2018b}.
 
-Previous work has suggested looking for SETI signals from interstellar ``beacons''
-\cite{benford2008}
-However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
+Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes.
+Such observing campaigns are expensive and difficult to obtain given the resource-limited nature of telescope allocation.
+Recent progress for a systematic ``technosignature'' search has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.
+
+The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When and where to look?}''. 
+While \citet{wright2018c} suggest the completeness of our search of the ``cosmic haystack'' at radio wavelengths is akin to the ratio of a swimming pool compared to the Earth's ocean's ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.
+However, new wide-field time domain surveys will enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing time resolved monitoring over many years.
+
+Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will also be transformative in enabling both reproducible and cost-efficient SETI.
+By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically.
+As new search algorithms are developed, these databases can be quickly re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.
+
+The data-driven spatial-temporal domain explored by these surveys necessitates a new generation of SETI approaches be developed.
+Technosignatures utilizing the spatial-temporal domain could include spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
+In this paper I provide an overview of the potential for new SETI studies based on time domain surveys.
+%In \S\ref{sec:surveys} I describe several relevant surveys that are promising data sources for deploying automated SETI algorithms.
+Additionally I describe (two?) examples of signals that could be implemented with these surveys, with the hope that the community will develop additional strategies.
+
+%Many have noted these new surveys can be used for SETI, such as looking for unusual transits (REFS), and they have the potential to produce viable, repeatable SETI searches from surveys if the techniques can be formed into survey or database queries and therefore run on nearly any survey, such as searching for ``impossible'' patterns or sequences in light curves (e.g.
+Davenport et al in prep)
+
+%\bf THE POINT: } in this new spatial-temporal domain, we should be looking for things that happen in space, such as over densities of events, or coordination of events like transits.
+
+%{\bf THIS PAPER NEEDS TO BE RETHOUGHT, IN LIGHT OF MORE SPECIFIC PAPER IDEA:}{\it SETI with ZTF}
+%
+%In the search for extraterrestrial intelligence (SETI), most approaches focus on direct detection of photons originating from transmission or waste energy.
+These searches occur over a range of wavelength regimes, and assume an extraterrestrial civilization will produce sufficient radiation to be detected apart from their parent star.
+While this may be feasible at radio wavelengths where Sun-like stars have lower emission, it is much more difficult to outshine the star in the optical regime where many present-day time domain surveys operate.
+%
+%
+%Blocking a significant fraction of light from a star may instead be much easier than outshining it \cite{arnold2005}.
+This has led to recent searches for transit-like signatures that may have artificial origins.
+For example, PAPER demonstrated the observable impact of transits by non-spherical objects \cite{arnold2005a}.
+%on the impact of artificial structures on transits. also this work on that:
+%\cite{wright2016}.
+%recent data from \Kepler \citep{borucki2010}, has found weird transit-like signals
+%\cite{boyajian2015}, 
+%which some have considered as SETI candidates. However, follow-up observations have yet to yield any confirming signatures, and instead this looks like a YSO w/ comets maybe \citep{lisse2015}.
+%
+%
+%Previous work has suggested looking for SETI signals from interstellar ``beacons''
+%\cite{benford2008}
+%However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
 This is a very expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
+%\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
+However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
+%\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
 
-optical seti
-http://arxiv.org/abs/astro-ph/0506758
+%optical seti
+%\citep{holder2005} optical pulse SETI 
+%
+%a review of Schelling points from \citet{wright2017}
+%lighthouses \citep{zuluaga2015}
+%active and passive SETI in coordination (both temporal and spatial) with galactic events like supernovae 
+%\citep{lemarchand1994}
+%coordinated times/places, also known as Schelling or Focal points \citep{schelling1960}, are the key to finding people in unknown time/space.
+%planets in Earth's transit zone \cite{heller2016} - maybe transits would be aligned for our viewing benefit?
+%
 
-Instead, much cheaper to block light, rather than shine it.
-\cite{arnold2005}
-on the feasibility of transits for use in SETI detection, and 
-\cite{arnold2005a}
-on the impact of artificial structures on transits. also this work on that:
-\cite{wright2016}.
-recent data from \Kepler \citep{borucki2010}, has found weird transit-like signals
-\cite{boyajian2015}, 
-which some have considered as SETI candidates. However, follow-up observations have yet to yield any confirming signatures, and instead this looks like a YSO w/ comets maybe \citep{lisse2015}.
-
-active and passive SETI in coordination (both temporal and spatial) with galactic events like supernovae 
-\citep{lemarchand1994}
-coordinated times/places, also known as Schelling or Focal points \citep{schelling1960}, are the key to finding people in unknown time/space.
-
-lighthouses
-\citep{zuluaga2015}
-
-planets in Earth's transit zone \cite{heller2016}
-
-\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
-However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
-
-\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
-
-a review of Schelling points from \citet{wright2017}
-
-In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization.
+%In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization.
 This system has the advantage of being potentially detectable in the near future via exoplanet searches.
 
+%%%%
+\section{A New Domain}
+we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
+
+bright facilities have excellent sky and time coverage:
+ASAS-SN \citep{kochanek2017}
+
+Evryscope \citep{law2015}
+
+starting to probe fainter sources becomes a trade off for area and time coverage.
+Zwicky Transient Facility \citep{bellm2014}
+for example surveys the sky every couple nights, but with only a single visit per source per night.
+
+space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
+The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
+and TESS \citep{tess}, as well as Gaia \citep{gaia} provide incredible precision light curves
+
+parameter space to search:
+- spatially coordinated events
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example: Coordinated Transits as ET Beacons}
+\section{Example 1: Coordinated Transits as ET Beacons}
 imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Searching for Correlated Transits in Kepler}
+%we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
+%
+%we find there is no correlation. Thats OK, but at least we searched
+%
+%\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
+%See also TCE histogram in \citet{twicken2016}
+%
+%
+%
+%\begin{figure}[]
+%\centering
+%\includegraphics[width=3in]{../figures/delta_per.png}
+%\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
+%median period differences are computed for bins of separation distance (orange points).
+%there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
+%\label{fig:2dcor}
+%\end{figure}
+%
+%
+%
+%\begin{figure}[]
+%\centering
+%\includegraphics[width=3in]{../figures/delta_per2.png}
+%\caption{correlation between transit periods and projected separations, as in Figure \ref{fig:2dcor}, but emphasizing the smallest separations in both axes.
+One outlier was identified in the similarity of Kepler 1292b and Kepler 1235b (black circle).
+However, these two systems are more than 600 pc separated in distance.}
+%\label{fig:2dcor2}
+%\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Prospects for Future Surveys}
@@
 \acknowledgments
-JRAD is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST-1501418.
+%JRAD is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST-1501418.
+%
+%This research has made use of the Exoplanet Orbit Database and the Exoplanet Data Explorer at exoplanets.org.
 
-\bibliography{/Users/james/research/references}
+JRAD acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
+The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
+
+\bibliography{/Users/james/Dropbox/references}
 

### 2018-11-28 code/haystack_fraction_calculator.py
first calculations of haystack for big surveys

@@ -1,0 +1,245 @@
+# based on code: https://github.com/shbhuk/SETI_haystack
+# published: https://doi.org/10.3847/1538-3881/aae099
+
+import numpy as np
+from astropy import units as u
+from astropy import constants as ac
+import sys,os
+import configparser
+
+def vol8d(home='', verbose=False, hayfile='Haystack_boundaries.txt', searchfile='Search_parameters.txt'):
+    '''
+    Read in the Haystack and search parameters from the config files.
+    The config (input) files must be stored in the same directory as this script.
+    Please use the default units for the observation parameters and haystack boundaries
+
+    The 9 dimensions are -
+    1) Sensitivity (m^2/W)
+    2) 3 Distance dimensions (m^3)
+    5) Transmitter BW (Hz)
+    6) Transmitter Frequency (Hz)
+    7) Repetition Time (sec)
+    8) Polarization
+
+    9) Modulation - not covered
+
+    '''
+
+    # allow users to specify their own "home directory" to read from
+    if len(home)<1:
+        home = sys.path[0]
+
+    # allow users to specify their own input files
+    haystack_boundary = os.path.join(home, hayfile)
+    search_parameters = os.path.join(home, searchfile)
+
+
+    haystack_config =  configparser.ConfigParser()
+    haystack_config.read(haystack_boundary)
+    search_config = configparser.ConfigParser()
+    search_config.read(search_parameters)
+
+
+    ##############################
+    # User observation  parameters
+    ##############################
+
+
+    # Total integration time for observation
+    total_scan = float(search_config['Search']['Time_on_Target']) * u.s # [seconds]
+
+    # Channel bandwidth (spectral resolution)
+    c = float(search_config['Search']['Channel_BW']) * u.Hz # [Hertz]
+
+    # Survey sensitivity
+    Survey_sensitivity = float(search_config['Search']['Survey_sensitivity']) * u.Jy # [Jansky]
+
[…]

### 2018-11-29 paper/ms.tex
a few more words added

@@
 The latest generation of synoptic time domain surveys  enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
-SETI methodology has not yet fully utilized such surveys, however, and the literature is in need of new search strategies.
-Here I describe the potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some new example SETI approaches.
+Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space than typical radio SETI investigations.
+Investigations with these surveys can also be conducted at a fraction of the cost, since they require only re-analysis of data already being gathered.
+However, SETI methodology has not widely utilized such surveys, and the literature is in need of new search algorithms that can incorporate signals from both the spatial and temporal domains.
+Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some new example SETI approaches that utilize the spatial-temporal domain
 \end{abstract}
@@
 
+In this paper I explore the potential of wide-field time domain surveys for conducting new technosignature searches.
+In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
 The data-driven spatial-temporal domain explored by these surveys necessitates a new generation of SETI approaches be developed.
 Technosignatures utilizing the spatial-temporal domain could include spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
-In this paper I provide an overview of the potential for new SETI studies based on time domain surveys.
-%In \S\ref{sec:surveys} I describe several relevant surveys that are promising data sources for deploying automated SETI algorithms.
-Additionally I describe (two?) examples of signals that could be implemented with these surveys, with the hope that the community will develop additional strategies.
+In \S\ref{sec:transit} I discuss one such signal, where transiting planets around many stars act as a spatially distributed beacon network.
+Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with optical time domain surveys.
 
@@
 %%%%
-\section{A New Domain}
+\section{Comparing Optical Surveys to Radio}
+\label{sec:haystack}
 we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example 1: Coordinated Transits as ET Beacons}
+\section{Transits as a Spatial-Temporal Beacon}
+%\section{Example 1: Coordinated Transits as ET Beacons}
+\label{sec:transit}
+
+\subsection{Coordinated Transit Networks}
 imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Searching for Correlated Transits in Kepler}
-%we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
-%
-%we find there is no correlation. Thats OK, but at least we searched
-%
-%\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
-%See also TCE histogram in \citet{twicken2016}
-%
-%
-%
-%\begin{figure}[]
-%\centering
-%\includegraphics[width=3in]{../figures/delta_per.png}
-%\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
-%median period differences are computed for bins of separation distance (orange points).
-%there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
-%\label{fig:2dcor}
-%\end{figure}
-%
-%
-%
+IDEA:
+could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
+How much more efficient would it be if we forced $i<10^\circ$, instead of $i<90^\circ$?
+
+for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
+by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
+
+PROBLEM:
+as \citet{forgan2017} note, an interstellar communications based on transits may only be stable for $\sim$10,000 years due to evolution of planetary orbits and relative motion between stars within the Milky Way.
+The beacon systems centered around the host system would have to be traveling as a moving group to maintain their respective trilateration signals, or have the transiting bodies actively over time.
+In other words, as the beacon systems moved relative to the ET host star due to differing Galactic orbits, the semi-major axis of the transiting bodies would have to be adjusted to compensate.
+
+%%%%%%%
+\subsection{Searching for Correlated Transits in Kepler}
+we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
+
+we find there is no correlation. Thats OK, but at least we searched
+
+\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
+See also TCE histogram in \citet{twicken2016}
+
+\begin{figure}[]
+\centering
+\includegraphics[width=3in]{../figures/delta_per.png}
+\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
+median period differences are computed for bins of separation distance (orange points).
+there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
+\label{fig:2dcor}
+\end{figure}
+
 %\begin{figure}[]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Prospects for Future Surveys}
+\section{Discussion}
+\label{sec:discussion}
+%
+%to get accurate census you need a complete time domain survey at some transit depth sensitivity out to some orbital period. going off our simulation, TESS might work for very short period planets
+%
+%TESS would be great for this in terms of spatial-temporal coverage. 
+%
+%LSST very good for finding SETI signals of this geometric style, but not ideal for events requiring such dedicating monitoring
+%
+%this kind of distributed network could be very efficient at broadcasting the presence of a civilization. dust clouds could create sufficient "transit" signals, while not affecting orbital dynamics
 
-to get accurate census you need a complete time domain survey at some transit depth sensitivity out to some orbital period. going off our simulation, TESS might work for very short period planets
+so far we have focused on the need for developing SETI approaches that utilize the new spatial-temporal domain.
+The data streams from these surveys also enable many other search strategies that could be implemented.
+these might include:
+VASCO-type events, stars appearing or disappearing. \citep{villarroel2016}
+	- Esp. stars appearing that do not have a Gaia DR2 detection, for example.
+	- need to account for novae, etc...
+	- stars disappearing will be very interesting
 
-TESS would be great for this in terms of spatial-temporal coverage.
+coordination/synchronization \citep{makovetskii1977,shostak2004}
+        - especially the SETI ellipsoid, a'la SN1987A \citep{lemarchand1994}
+        - could use galactic novae to provide time/place
+        - %https://asd.gsfc.nasa.gov/Koji.Mukai/novae/novae.html (new Novae)
+        - %http://www.cbat.eps.harvard.edu/nova_list.html historic Novae (till 2010)
 
-LSST very good for finding SETI signals of this geometric style, but not ideal for events requiring such dedicating monitoring
+unnaturals patterns of otherwise normal events
+	- look for e.g. flares or novae that recur w/ prime or fibonacci sequences
+	- see forthcoming paper
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-this kind of distributed network could be very efficient at broadcasting the presence of a civilization. dust clouds could create sufficient "transit" signals, while not affecting orbital dynamics
+Surveys like ZTF (and soon LSST) publish real-time alerts of variability from sources. These alert streams could be monitored in real-time for signals as well:
+alerts from the stars within the ``restricted Earth Transit Zone'' \citep{heller2016}
+	- anytime we can see them w/ ZTF
+	- especially at our opposition (i.e. when we transit)
+	- with Gaia we can track stars that are leaving the ETZ as well, but might 
+	have seen us transit previously (or by the time the signal reaches us)
 
-IDEA:
-could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
-How much more efficient would it be if we forced $i<10^\circ$, instead of $i<90^\circ$?
+alerts from known exoplanet host systems, especially coinciding with any known transit conjunction (mid-transit), 
+	- allow some tolerance window to account for TTVs
 
-for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
-by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
+The cost of deploying such SETI is remarkably low, since they can ``piggyback'' on databases and search tools being developed for the primary science drivers of each survey (e.g. transients for ZTF).
+
+%Lastly, we should resist the impulse to try and develop every possible search algorithm to mine this wealth of data for technosignatures.
+Instead, immediate progress can be made in SETI by deploying robust algorithms to search for specific signals
 
@@
 \acknowledgments
-%JRAD is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST-1501418.
-%
-%This research has made use of the Exoplanet Orbit Database and the Exoplanet Data Explorer at exoplanets.org.
+
+The author thanks Nicholas M.
+Law and Jason T.
+Wright for their helpful ongoing conversations in developing this idea, and Andrew P.
+V.
+Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit.
 

### 2019-03-21 paper/ms.tex
add section to write

@@
 \author{James. R. A. Davenport}
-%\affiliation{Department of Physics \& Astronomy, Western Washington University, 516 High St., Bellingham, WA 98225, USA}
-%\altaffiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
 \affiliation{Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
@@
 The latest generation of synoptic time domain surveys  enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
-Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space than typical radio SETI investigations.
+Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space volume than many radio SETI investigations.
+Small-aperture, high cadence surveys like Evryscope can be comparable in their Cosmic Haystack volume completeness to deeper surveys including LSST.
 Investigations with these surveys can also be conducted at a fraction of the cost, since they require only re-analysis of data already being gathered.
@@
 
-In this paper I explore the potential of wide-field time domain surveys for conducting new technosignature searches.
+In this paper I explore the potential of wide-field time domain surveys for conducting a new generation of technosignature searches. 
+In \S\ref{sec:method} I describe the general principle behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
 In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
-The data-driven spatial-temporal domain explored by these surveys necessitates a new generation of SETI approaches be developed.
-Technosignatures utilizing the spatial-temporal domain could include spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
-In \S\ref{sec:transit} I discuss one such signal, where transiting planets around many stars act as a spatially distributed beacon network.
-Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with optical time domain surveys.
+The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon network.
+Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.
 
@@
 %%%%
-\section{Comparing Optical Surveys to Radio}
+\section{Optical Survey SETI}
+\label{sec:method}
+METHOD: what is the big idea, in detail
+
+include mantras of "worth doing", "do it almost for free", "portable, do it everywhere", and "outliers are nature papers"
+
+%%%%
+\section{Comparing Optical Surveys to Radio Searches}
 \label{sec:haystack}
@@
 
+\begin{figure*}[]
+\centering
+\includegraphics[width=3.5in]{../figures/optical_haystack}
+\caption{
+comparison of the 8D ``Cosmic Haystack'' volume defined by \citet{wright2018c} using four optical surveys with varying designs.
+}
+\label{fig:hay}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Transits as a Spatial-Temporal Beacon}
+\section{Example: a Spatial-Temporal Transit Beacon}
 %\section{Example 1: Coordinated Transits as ET Beacons}

### 2019-04-09 paper/ms.tex
add a paragraph monday

@@
-\documentclass[modern]{aastex62}
+\documentclass[twocolumn]{aastex62}
 \bibliographystyle{aasjournal}
@@
 Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes.
-Such observing campaigns are expensive and difficult to obtain given the resource-limited nature of telescope allocation.
-Recent progress for a systematic ``technosignature'' search has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.
+Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation.
+Recent progress for systematic ``technosignature'' searchs has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.
 
@@
 
-Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will also be transformative in enabling both reproducible and cost-efficient SETI.
+Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI.
 By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically.
@@
 
-In this paper I explore the potential of wide-field time domain surveys for conducting a new generation of technosignature searches. 
-In \S\ref{sec:method} I describe the general principle behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
+In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches. 
+In \S\ref{sec:method} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
 In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
@@
 Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.
-
-%Many have noted these new surveys can be used for SETI, such as looking for unusual transits (REFS), and they have the potential to produce viable, repeatable SETI searches from surveys if the techniques can be formed into survey or database queries and therefore run on nearly any survey, such as searching for ``impossible'' patterns or sequences in light curves (e.g.
-Davenport et al in prep)
 
@@
 
-%In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization.
-This system has the advantage of being potentially detectable in the near future via exoplanet searches.
-
 %%%%
@@
 \label{sec:method}
-METHOD: what is the big idea, in detail
 
-include mantras of "worth doing", "do it almost for free", "portable, do it everywhere", and "outliers are nature papers"
+Wide-field imaging surveys, and their associated infrastructure, are an ideal data source for carrying out technosignature searches that augment efforts in other wavelengths or with different observing strategies.
+By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look.
+Increasingly, multiple imaging surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g.
+imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events..
+This also results in survey data providing some ``ground truth'' about the nature of most known objects in our Galaxy (e.g.
+their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} we're searching for technosignatures.
+
+% Reproducible science
+The large archives of data produced from these surveys become valuable legacies that enable additional science long after data acquisition has concluded.
+
+Studies for new phenomena in time-domain astronomy can therefore be carried out both in the future with new observations and in the past by re-analyzing survey archives.
+
+%somrthing about sky surveys \citep{djorgovski2013}
+
+%include mantras of "worth doing", "do it almost for free", "reproducible", "portable, do it everywhere", and "outliers are nature papers"
+
+large ground-based surveys, such as LSST, are a highly ranked component of advancement in our current landscape something something
+\citep{astro2010}.
+
+real time tools being developed for LSST \citep{schwamb2019}
+
+RNAAS about tools to model technosignatures from radio surveys
+\citep{forgan2019}
+
+\citep{lacki2019} looking for specular reflections of interplanetary objects, particularly in big new surveys
 

### 2019-04-16 paper/ms.tex
section 2 written thru

@@
 Small-aperture, high cadence surveys like Evryscope can be comparable in their Cosmic Haystack volume completeness to deeper surveys including LSST.
-Investigations with these surveys can also be conducted at a fraction of the cost, since they require only re-analysis of data already being gathered.
+Investigations with these surveys can also be conducted at a fraction of the cost, since they require analysis of data already being gathered.
 However, SETI methodology has not widely utilized such surveys, and the literature is in need of new search algorithms that can incorporate signals from both the spatial and temporal domains.
@@
 
-%%%%
-\section{Optical Survey SETI}
+% include mantras of "worth doing", "do it almost for free", "reproducible", "portable, do it everywhere", and "outliers are nature papers"
+
+%%%%%%%%%%%%%%%%
+\section{Why Optical Survey SETI?}
 \label{sec:method}
 
+% The first SETI challenges: WHEN, WHERE, WHAT to search
 Wide-field imaging surveys, and their associated infrastructure, are an ideal data source for carrying out technosignature searches that augment efforts in other wavelengths or with different observing strategies.
 By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look.
-Increasingly, multiple imaging surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g.
+Increasingly, multiple surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g.
 imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events..
 This also results in survey data providing some ``ground truth'' about the nature of most known objects in our Galaxy (e.g.
-their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} we're searching for technosignatures.
+their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} objects we're studying for SETI.
 
-% Reproducible science
-The large archives of data produced from these surveys become valuable legacies that enable additional science long after data acquisition has concluded.
+% next: HOW to search (can compare approaches, and make them reproducible)
+Optical surveys are also an excellent platform for exploring {\it how} to conduct technosignature searches.
+The large archives of data produced from these surveys become valuable legacies that enable science long after data acquisition has concluded.
+Searches for new phenomena in time-domain astronomy can be carried out both in the future with new observations and in the past by re-analyzing survey archives.
+Using survey archives for SETI as opposed to targeted observations allows us to compare the results of competing algorithms or strategies, and enables reproducible science.
+Search algorithms are also portable, and ideally can be used to carry out SETI on any appropriately structured large survey.
+As future survey data is gathered, new classes of technosignature signals may be detectable that rely on decades worth of data. 
+Since any candidate technosignature detection will (or should) be the subject of great scrutiny, making the data and algorithms open and reproducible will be critical for comparison and independent validation.
 
-Studies for new phenomena in time-domain astronomy can therefore be carried out both in the future with new observations and in the past by re-analyzing survey archives.
+% DO IT FOR "FREE"
+Large surveys, particularly ground-based efforts like LSST, are a highly prioritized component of the current and future development landscape in astronomy \citep{astro2010}.
+As such, great efforts are being made to develop survey tools and technologies to enable science from these missions, such as new database technologies \citep{juric2012}.
+These developments are especially needed to utilize the new forms of data being made available, including real-time data analysis \citep[e.g.][]{schwamb2019} and broadcasting real-time alerts for events \citep{patterson2019}.
+The infrastructure needed to conduct technosignature searches with these large optical surveys is therefore already being developed.
+As with many areas of astrophysics in the ``survey era'', the biggest restriction to SETI with large surveys is therefore the availability of supported researchers mine the data.
 
-%somrthing about sky surveys \citep{djorgovski2013}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Potential Strategies}
 
-%include mantras of "worth doing", "do it almost for free", "reproducible", "portable, do it everywhere", and "outliers are nature papers"
-
-large ground-based surveys, such as LSST, are a highly ranked component of advancement in our current landscape something something
-\citep{astro2010}.
-
-real time tools being developed for LSST \citep{schwamb2019}
-
-RNAAS about tools to model technosignatures from radio surveys
-\citep{forgan2019}
+SETI with large surveys falls into two basic approaches: 1) classify every object or phenomena in the data, and any remaining outliers are technosignature candidates, or 2) search for specific ``impossible'' or unusual types of signals in the data, which I advocate for in this work.
+While the latter strategy does not guarantee a complete search for all SETI signals in the data, it is technically feasible for almost any time domain survey.
 
@@
 
-%%%%
-\section{Comparing Optical Surveys to Radio Searches}
+several kinds of signals have been suggested that may be viable. Here I list a few general classes of technosignature signals that may be well suited for developing detection algorithms around:
+
+\begin{itemize}
+\item unusual variability -- either on short timescales (e.g. Boyajian's Star) or long timescales (e.g. VASCO)
+\item spatial correlations of events or phenomena -- e.g.
+coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse'' \citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
+\item spatial density or distribution -- e.g. an over density of phenomena, such as within the Earth Transit Zone \citep{heller2016}
+\item unnatural patterns, particularly of otherwise normal variability -- e.g. a prime or fibonacci sequence
+\end{itemize}
+
+%%%%%%%%%%%%%%%%
+\section{Comparing Optical Surveys\\ to Radio Searches}
 \label{sec:haystack}
+
 we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
@@
 
-Evryscope \citep{law2015}
+Evryscope \citep{law2015} is amazing
 
@@
 The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
-and TESS \citep{tess}, as well as Gaia \citep{gaia} provide incredible precision light curves
+and TESS \citep{tess}, as well as Gaia \citep{gaia} provide incredible precision light curves for millions of stars.
 
-parameter space to search:
-- spatially coordinated events
+RNAAS about tools to model technosignatures from radio surveys
+\citep{forgan2019}
 
-\begin{figure*}[]
+\begin{figure}[]
 \centering
@@
 \label{fig:hay}
-\end{figure*}
+\end{figure}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example: a Spatial-Temporal Transit Beacon}
+\section{Example: Spatial-Temporal\\ Transit Beacons}
 %\section{Example 1: Coordinated Transits as ET Beacons}

### 2019-04-16 paper/ms.tex
start writing section 3

@@
 
-In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches. 
-In \S\ref{sec:method} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
+In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches, starting with an overview of the motivation for doing SETI with optical surveys in \S\ref{sec:method}.
+In \S\ref{sec:strategy} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
 In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
@@
 \section{Potential Strategies}
+\label{sec:strategy}
 
@@
 
-\citep{lacki2019} looking for specular reflections of interplanetary objects, particularly in big new surveys
+Several types of signals have been suggested that may be viable for use with optical surveys.
+For example, \citet{villarroel2016} searched two wide field, single-epoch optical surveys for objects ``disappearing''  over decades timescales.
+\citet{lacki2019} outline a method to search for specular reflections of interplanetary objects, particularly in wide field surveys.
+\citet{arnold2005} describe the the observable signature of transits from ostensibly artificial objects, which could be searched for with exoplanet transit surveys.
 
-several kinds of signals have been suggested that may be viable. Here I list a few general classes of technosignature signals that may be well suited for developing detection algorithms around:
+Here I list a few general classes of technosignature signals that may be well suited for developing new detection algorithms around:
 
 \begin{itemize}
-\item unusual variability -- either on short timescales (e.g. Boyajian's Star) or long timescales (e.g. VASCO)
+\item unusual variability profiles or statistical distributions -- either on short timescales (e.g. Boyajian's Star) or long timescales (e.g. VASCO)
 \item spatial correlations of events or phenomena -- e.g.
@@
 \item spatial density or distribution -- e.g. an over density of phenomena, such as within the Earth Transit Zone \citep{heller2016}
-\item unnatural patterns, particularly of otherwise normal variability -- e.g. a prime or fibonacci sequence
+\item unnatural patterns, particularly patterns of otherwise normal variability -- e.g. a prime or fibonacci sequence
 \end{itemize}

### 2019-04-17 paper/ms.tex
section 3 written thru

@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{SETI in the Optical Spatial-Temporal Domain}
+\title{SETI in the Spatial-Temporal Survey Domain}
 
-\shorttitle{Spatial-Temporal SETI}
+\shorttitle{Survey SETI}
 \shortauthors{Davenport}
@@
 
-The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When and where to look?}''. 
+The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When, where, and how to look?}''. 
 While \citet{wright2018c} suggest the completeness of our search of the ``cosmic haystack'' at radio wavelengths is akin to the ratio of a swimming pool compared to the Earth's ocean's ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.
@@
 \citet{lacki2019} outline a method to search for specular reflections of interplanetary objects, particularly in wide field surveys.
-\citet{arnold2005} describe the the observable signature of transits from ostensibly artificial objects, which could be searched for with exoplanet transit surveys.
+\citet{arnold2005} describe the the observable signature of transits from artificial objects, which could be searched for with exoplanet transit surveys.
+\citet{kipping2016} find that exoplanet transits could be hidden or ``cloaked'' to an observer by using directed laser emission.
 
-Here I list a few general classes of technosignature signals that may be well suited for developing new detection algorithms around:
+Here I describe a few general classes of technosignature signals that may be well suited to developing new detection algorithms for spatial-temporal surveys:
 
-\begin{itemize}
-\item unusual variability profiles or statistical distributions -- either on short timescales (e.g. Boyajian's Star) or long timescales (e.g. VASCO)
-\item spatial correlations of events or phenomena -- e.g.
+\begin{enumerate}
+\item {\bf Unusual variability profiles or statistical distributions of fluxes }-- either on short timescales such as Boyajian's Star \citep{boyajian2015}, or long timescales such as disappearing stars \citep{villarroel2016} or occasionally missing transits \citep{kipping2016}
+%
+\item {\bf Spatial correlations of events or phenomena }-- e.g.
 coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse'' \citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
-\item spatial density or distribution -- e.g. an over density of phenomena, such as within the Earth Transit Zone \citep{heller2016}
-\item unnatural patterns, particularly patterns of otherwise normal variability -- e.g. a prime or fibonacci sequence
-\end{itemize}
+%
+\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the Earth Transit Zone \citep{heller2016}
+%
+\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability, such as broadcasting a prime or fibonacci sequence of transits \citep{wright2016}.
+%
+\end{enumerate}
 

### 2019-04-17 paper/ms.tex
start writing Section 4

@@
 
-we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
+\citet{wright2018c} have provided a framework to quantify the volume of observable parameter space covered by various SETI projects -- dubbed the ``Cosmic Haystack''.
+The properties for several major radio SETI projects yield a Haystack volume coverage of $10^{-20}$ to $10^{-18}$, or as \citet{wright2018c} illustrate, akin to a drinking glass of water compared to the volume of Earth's oceans.
+While this tool was developed with radio surveys in mind, it is broadly applicable to surveys at other wavelengths that are not explicitly designed for SETI \citep[e.g.
+see][]{forgan2019}.
 
-bright facilities have excellent sky and time coverage:
-ASAS-SN \citep{kochanek2017}
+While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is useful to quantitatively compare their capability to traditional searches in the radio.
+For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Cosmic Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
+%we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
 
@@
 
-space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
+%space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
 The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
 and TESS \citep{tess}, as well as Gaia \citep{gaia} provide incredible precision light curves for millions of stars.
-
-RNAAS about tools to model technosignatures from radio surveys
-\citep{forgan2019}
 
@@
 \caption{
-comparison of the 8D ``Cosmic Haystack'' volume defined by \citet{wright2018c} using four optical surveys with varying designs. 
+comparison of the 8D ``Cosmic Haystack'' volume defined by \citet{wright2018c} using five optical surveys with varying designs. 
 }

### 2019-04-18 paper/ms.tex
section 4 written thru

@@
 
-While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is useful to quantitatively compare their capability to traditional searches in the radio.
+While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is important to quantitatively compare them to traditional searches in the radio.
 For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Cosmic Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
+Since these terms all act as scalers in the Haystack volume integration, surveys with different properties (e.g.
+few deep exposures versus many shallow exposures) can have comparably large Haystack coverage, filling out very different portions of the search volume.
+For maximum impact, SETI should therefore be conducted on many types of surveys.
 %we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
 
-Evryscope \citep{law2015} is amazing
-
-starting to probe fainter sources becomes a trade off for area and time coverage.
-Zwicky Transient Facility \citep{bellm2014}
-for example surveys the sky every couple nights, but with only a single visit per source per night.
-
-%space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
-The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
-and TESS \citep{tess}, as well as Gaia \citep{gaia} provide incredible precision light curves for millions of stars.
+As a demonstration, I have computed the Cosmic Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}.
+This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}.
+These include space-based and traditional ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design.
+%TESS and Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) provide
+I have assumed sensitivity ($S$) is the published per-visit photometric depth for each survey.
+The total integration time per target ($T$) was computed as the average number of visits per target in the survey multiplied by the per-visit exposure time.
+The same power of transmitter is assumed as in \citet{wright2018c}.
+The code to generate the figures in this manuscript is online.\footnote{\url{https://github.com/TheAstroFactory/survey_seti}}
 
@@
 \caption{
-comparison of the 8D ``Cosmic Haystack'' volume defined by \citet{wright2018c} using five optical surveys with varying designs. 
+comparison of the 8D ``Cosmic Haystack'' volume fraction defined by \citet{wright2018c} using five optical surveys with varying designs. 
 }
@@
 \end{figure}
+
+Typical values for the Haystack fraction for radio SETI programs as presented in \citet{wright2018c}  are in the range of $\log V_{Haystack} = $ 20 --17.
+However, these programs are mostly targeted observations of nearby stars.
+The largest program featured in \citet{wright2018c}, by \citet{tingay2018}, had $\Omega=400$ sq deg and $T=10800$ seconds, and yielded $\log V_{Haystack} =  -17.36$.
+
+By comparison, the wide field optical surveys I have featured all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction.
+Due to its very wide field coverage, and near continuous monitoring each night, Evryscope ranks as the best optical survey for technosignature searches of those surveyed here.
+The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Cosmic Haystack volume with as many different survey ``shapes'' as possible.
+
+While \citet{wright2018c} likened their Haystack fraction to searching for fish in a drinking glass of water compared to the Earth's oceans, our results are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}.
+While optical surveys still only cover a tiny fraction of the total volume of parameter space, it is a substantial improvement.
+
+%Evryscope \citep{law2015} is amazing
+%
+%starting to probe fainter sources becomes a trade off for area and time coverage.
+%Zwicky Transient Facility 
+%for example surveys the sky every couple nights, but with only a single visit per source per night.
+%
+
+%space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
+The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
+%, as well as Gaia \citep{gaia} provide incredible precision light curves for millions of stars.
 

### 2019-04-18 paper/ms.tex
add caption

@@
 \caption{
-comparison of the 8D ``Cosmic Haystack'' volume fraction defined by \citet{wright2018c} using five optical surveys with varying designs. 
-}
+Comparison of the 8-D ``Cosmic Haystack'' SETI search volume fraction, defined by \citet{wright2018c}, computed using five optical surveys with varying designs.
+The Haystack fraction covered by these optical surveys is 1-2 dex larger than typical SETI programs conducted in the radio.
+Evryscope \citet{law2015} is the best survey considered here for SETI work, narrowly beating LSST due to its wide field of view and very dense light curves.}
 \label{fig:hay}

### 2019-04-19 paper/ms.tex
a couple words today

@@
 These include space-based and traditional ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design.
-%TESS and Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) provide
 I have assumed sensitivity ($S$) is the published per-visit photometric depth for each survey.
@@
 By comparison, the wide field optical surveys I have featured all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction.
-Due to its very wide field coverage, and near continuous monitoring each night, Evryscope ranks as the best optical survey for technosignature searches of those surveyed here.
+Due to its very wide field coverage, and near continuous monitoring each night, Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) ranks as the best optical survey for technosignature searches of those surveyed here.
 The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Cosmic Haystack volume with as many different survey ``shapes'' as possible.
 
-While \citet{wright2018c} likened their Haystack fraction to searching for fish in a drinking glass of water compared to the Earth's oceans, our results are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}.
+\citet{wright2018c} likened their Haystack fraction to searching for fish in a drinking glass of water compared to the Earth's oceans,.
+Our results for SETI with optical surveys are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}.
 While optical surveys still only cover a tiny fraction of the total volume of parameter space, it is a substantial improvement.
-
-%Evryscope \citep{law2015} is amazing
-%
-%starting to probe fainter sources becomes a trade off for area and time coverage.
-%Zwicky Transient Facility 
-%for example surveys the sky every couple nights, but with only a single visit per source per night.
-%
-
-%space-based platforms designed for exoplanet searches provide very high quality light curves over months to years timescales.
-The \Kepler and K2 missions \citep{borucki2010,howell2014} produced 30-minute cadence light curves for $\sim$500,000 stars, with baselines ranging from $\sim$3 months up to 4 years.
-%, as well as Gaia \citep{gaia} provide incredible precision light curves for millions of stars.
 
@@
 
+To illustrate the potential for possible technosignature searches in the spatial-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets.
+
+%%%%%%%
+\subsection{Correlated Transits in Kepler}
+we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
+
+we find there is no correlation. Thats OK, but at least we searched
+
+\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
+See also TCE histogram in \citet{twicken2016}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3in]{../figures/delta_per.png}
+\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
+median period differences are computed for bins of separation distance (orange points).
+there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
+\label{fig:2dcor}
+\end{figure}
+
+%%%%%%
 \subsection{Coordinated Transit Networks}
@@
 
-\begin{figure}[]
-\centering
-\includegraphics[width=3in]{../figures/recov_fraction.pdf}
-\caption{fraction of hot Jupiter-like systems recovered in 1000 realizations of our 1000-star simulation.
+%\begin{figure}[]
+%\centering
+%\includegraphics[width=3in]{../figures/recov_fraction.pdf}
+%\caption{fraction of hot Jupiter-like systems recovered in 1000 realizations of our 1000-star simulation.
 This curve is dependent on the ratio of the occulter-to-star size, and the orbital period range sampled.
-The average from our model is $10\pm 3$ systems recovered for these parameters.}
-\label{fig:recov}
-\end{figure}
+%The average from our model is $10\pm 3$ systems recovered for these parameters.}
+%\label{fig:recov}
+%\end{figure}
 
@@
 In other words, as the beacon systems moved relative to the ET host star due to differing Galactic orbits, the semi-major axis of the transiting bodies would have to be adjusted to compensate.
-
-%%%%%%%
-\subsection{Searching for Correlated Transits in Kepler}
-we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
-
-we find there is no correlation. Thats OK, but at least we searched
-
-\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
-See also TCE histogram in \citet{twicken2016}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=3in]{../figures/delta_per.png}
-\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
-median period differences are computed for bins of separation distance (orange points).
-there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
-\label{fig:2dcor}
-\end{figure}
-
-%\begin{figure}[]
-%\centering
-%\includegraphics[width=3in]{../figures/delta_per2.png}
-%\caption{correlation between transit periods and projected separations, as in Figure \ref{fig:2dcor}, but emphasizing the smallest separations in both axes.
-One outlier was identified in the similarity of Kepler 1292b and Kepler 1235b (black circle).
-However, these two systems are more than 600 pc separated in distance.}
-%\label{fig:2dcor2}
-%\end{figure}
 

### 2019-04-23 paper/ms.tex
starting to restructure the example section

@@
 %%%%%%%
-\subsection{Correlated Transits in Kepler}
+%\subsection{Correlated Transits in Kepler}
 we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
@@
 
-\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
-See also TCE histogram in \citet{twicken2016}
+%\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
+%See also TCE histogram in \citet{twicken2016}
 
-\begin{figure}[!t]
+\begin{figure}[!h]
 \centering
@@
 %%%%%% 
-\subsection{Coordinated Transit Networks}
-imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
-
-give the detailed example i have thought up.
-
-becons placed at distances
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=2.75in]{../figures/dist_per.pdf}
-\includegraphics[width=2.75in]{../figures/sky_per.pdf}
-\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
-\label{fig:2d}
-\end{figure*}
-
-for clarity, this is what the ideal signal might look like on the sky
-
-%%%
-
-this type of beacon network is advantageous because it points directly back towards their home. only a few systems actually need to be transiting from any given line of sight.
-
-NOW THE FULL SIMULATION...
-- computation to do: if had 100 beacons, each placed at random orbital alignment, in 3d sphere around home system.
-- assume G stars, 
-- odds of observing transit of a fixed sized object versus orbital distance... goes down. need that plot to figure out probability. 
-- assume ET places beacons with uniform RADIAL density in 3d space out to some maximum distance (even \# of systems as function of radial distance) in bins of 10pc out to 100 pc (i.e.
+%    > shelve this idea for now, maybe another paper.... 1 demo is enough for this paper, 
+%    and it should be actually computed/worked out.
+%%%%%% 
+%\subsection{Coordinated Transit Networks}
+%imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
+%
+%give the detailed example i have thought up.
+%
+%becons placed at distances
+%
+%\begin{figure*}[]
+%\centering
+%\includegraphics[width=2.75in]{../figures/dist_per.pdf}
+%\includegraphics[width=2.75in]{../figures/sky_per.pdf}
+%\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
+%\label{fig:2d}
+%\end{figure*}
+%
+%
+%for clarity, this is what the ideal signal might look like on the sky
+%
+%
+%this type of beacon network is advantageous because it points directly back towards their home. only a few systems actually need to be transiting from any given line of sight.
+%
+%NOW THE FULL SIMULATION...
+%- computation to do: if had 100 beacons, each placed at random orbital alignment, in 3d sphere around home system.
+%- assume G stars, 
+%- odds of observing transit of a fixed sized object versus orbital distance... goes down. need that plot to figure out probability. 
+%- assume ET places beacons with uniform RADIAL density in 3d space out to some maximum distance (even \# of systems as function of radial distance) in bins of 10pc out to 100 pc (i.e.
 10 beacons in each 10pc bin)
-- orbital period is exact for each system, no bins of period
-- do Monte Carlo sim with these parameters to figure out how many transiting systems we'd observe
-
----- make the plot for one MC realization of RA,Dec.... open circles for systems with no observed transits, colored for transits
-
-\begin{figure}[]
-\centering
-\includegraphics[width=3in]{../figures/3d_model.pdf}
-\caption{the model in 3 dimensions projected into the sky plane.
+%- orbital period is exact for each system, no bins of period
+%- do Monte Carlo sim with these parameters to figure out how many transiting systems we'd observe 
+%
+%---- make the plot for one MC realization of RA,Dec.... open circles for systems with no observed transits, colored for transits
+%
+%\begin{figure}[]
+%\centering
+%\includegraphics[width=3in]{../figures/3d_model.pdf}
+%\caption{the model in 3 dimensions projected into the sky plane.
 1000 simulated systems that span the galactic distance of 5-50 pc (blue) with 10 recovered transits highlighted (red).
 these would be hot Jupiters at the short P end.}
-\label{fig:3d}
-\end{figure}
-
-%\begin{figure}[]
-%\centering
-%\includegraphics[width=3in]{../figures/recov_fraction.pdf}
-%\caption{fraction of hot Jupiter-like systems recovered in 1000 realizations of our 1000-star simulation.
-This curve is dependent on the ratio of the occulter-to-star size, and the orbital period range sampled.
-%The average from our model is $10\pm 3$ systems recovered for these parameters.}
-%\label{fig:recov}
+%\label{fig:3d}
 %\end{figure}
-
-IDEA:
-could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
+%
+%
+%IDEA:
+%could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
 How much more efficient would it be if we forced $i<10^\circ$, instead of $i<90^\circ$?
-
-for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
-by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
-
-PROBLEM:
-as \citet{forgan2017} note, an interstellar communications based on transits may only be stable for $\sim$10,000 years due to evolution of planetary orbits and relative motion between stars within the Milky Way.
+%
+%for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
+%by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
+%
+%
+%PROBLEM:
+%as \citet{forgan2017} note, an interstellar communications based on transits may only be stable for $\sim$10,000 years due to evolution of planetary orbits and relative motion between stars within the Milky Way.
 The beacon systems centered around the host system would have to be traveling as a moving group to maintain their respective trilateration signals, or have the transiting bodies actively over time.
 In other words, as the beacon systems moved relative to the ET host star due to differing Galactic orbits, the semi-major axis of the transiting bodies would have to be adjusted to compensate.
+%
 

### 2019-04-23 paper/ms.tex
better idea for examples

@@
 
-%%%%%%%
-%\subsection{Correlated Transits in Kepler}
-we compute the 2-point correlation between all transits in the Kepler sample in 2D space (radius)
+trivial example: two point correlation of exoplanets, not substantively different from field stars, so not a total engineering of the Kepler field. OK
 
-we find there is no correlation. Thats OK, but at least we searched
+\begin{figure}[!h]
+\centering
+\includegraphics[width=3in]{../figures/2ptcf.pdf}
+\caption{two point correlation function of transiting exoplanets versus all stars in the kepler field.
+No obvious signs of engineering of the exoplanet population as a whole.
+exoplanets are actually slightly less correlated on small size scales than stars in general, which may be an artifact of detected planets being around brighter stars in the sample (i.e.
+having lower total density).
+While overly simple, this example demonstrates one possible test that could be done using existing and future exoplanet surveys.}
+\label{fig:2pt}
+\end{figure}
+
+more interesting example: how to stand out in exoplanet signals?
+if the simplest signal is just to indicate the presence of ET, then Q is how to stand out from the field?
+one idea is a spatial correlation of some exoplanet property, such as radius, eclipse depth, or orbital period.
+here i propose we search for spatial clusters of exoplanets with unusually similar orbital periods.
+
+to make an ideal transmitter of such a signal, the Q's are: what period to use, and how many systems need to be detected as similar to be confident something is odd about them?
+the first Q a civilization might consider either periods that are unusual (long P i think, look at population histogram).
+But to build such a beacon, might be efficient to balance with detectability since several systems must be detected to stand out, so short orbital period is probably best (easiest to detect).
+
+the second Q is akin to a loaded dice problem (i.e.
+how many rolls do you need to tell a die isnt rolling fair).
+The odds of a single random transiting system being a given period is just a random draw from the underlying population, so the height of a bin in the normalized distribution.
+The odds of N draws is equivalent to multiplying the normalized distribution by itself N times.
+So looking at a new Figure i'm going to make, we need 4 (?) systems to be within the same bin to be highly unusual...
+use this to drive choice of N, and for a given result (if positive) estimate the probability
+
+so for every planet in Kepler, I'll measure the.... stddev? sum of sq differences? some metric of "sameness"
+of the 4 (?) nearest exoplanets.
+then the Q is what's the  99\% outlier for the sameness distribution?
+choosing to do this simple numerical approach since the data are noisy, no way to exactly have ``same'' period....
+
+probably find no candidates, but that'll be AOK since the odds of 4 draws having exact same period should be very low.
 
@@
 
-\begin{figure}[!h]
-\centering
-\includegraphics[width=3in]{../figures/delta_per.png}
-\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
-median period differences are computed for bins of separation distance (orange points).
-there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
-\label{fig:2dcor}
-\end{figure}
+%
+%\begin{figure}[!h]
+%\centering
+%\includegraphics[width=3in]{../figures/delta_per.png}
+%\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
+%median period differences are computed for bins of separation distance (orange points).
+%there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
+%\label{fig:2dcor}
+%\end{figure}
 

### 2019-04-24 paper/ms.tex
into Section 5 now

@@
 %
-\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the Earth Transit Zone \citep{heller2016}
+\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the Earth Transit Zone \citep{heller2016}, or in spatial clusters
 %
-\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability, such as broadcasting a prime or fibonacci sequence of transits \citep{wright2016}.
+\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability, such as broadcasting a prime or fibonacci sequence of transits \citep{arnold2005,wright2016}.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example: Spatial-Temporal\\ Transit Beacons}
+\section{Example: Distribution of Transiting Exoplanets in the {\it Kepler} Field}
 %\section{Example 1: Coordinated Transits as ET Beacons}
@@
 To illustrate the potential for possible technosignature searches in the spatial-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets.
+Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or TESS, and they represent a natural ``Schelling point'' to attract attention with \citep{wright2017}.
+Creating such a beacon requires either generating a transiting system with unusual properties or patterns \citep[e.g.][]{arnold2005,kipping2016,forgan2017}, or as outlined in \S\ref{sec:strategy} an unusual {\it distribution} of transiting systems.
+The two examples outlined here explore the latter scenario, though I note the former is also well suited to automated approaches that would make use of future surveys and should be developed further.
 
-trivial example: two point correlation of exoplanets, not substantively different from field stars, so not a total engineering of the Kepler field. OK
+%__________
+\subsection{Two-Point Correlation Function}
 
-\begin{figure}[!h]
+\citep{landy1993}
+
+trivial example: two point correlation of exoplanets, Figure \ref{fig:2pt}
+ not substantively different from field stars, so not a total engineering of the Kepler field.
+OK however note the selection function of these two populations is not exactly the same, so this can be a tricky metric to  use.
+considerable effort to ensure well understood sample and biases from non-uniform spatial selection, even in ``simple'' scenarios of galaxy populations in well-studied surveys \citep[e.g.][]{wang2013}.
+also requires large samples (like with simple spatial overdensities) to determine if something is unusual.
+This makes it an expensive beacon to create
+
+\begin{figure}[!t]
 \centering
@@
 
+%__________
+\subsection{Clusters of Similarity}
 more interesting example: how to stand out in exoplanet signals?
@@
 probably find no candidates, but that'll be AOK since the odds of 4 draws having exact same period should be very low.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3in]{../figures/Kplanets.pdf}
+\caption{Spread in orbital periods for KNN clusters of exoplanet systems in \Kepler versus the average angular separation of the cluster.
+For $K=3$, 5 clusters have a spread in their periods below 6\% (red points).
+These KNN clusters have typical spatial separations of $\sim$0.1 deg.}
+\label{fig:knn}
+\end{figure}
 

### 2019-04-25 paper/ms.tex
clean comments from removed text out
wednesday progress

@@
 
-%Here I propose one such SETI approach, which utilizes exoplanet-like transit signals coordinated around multiple stars to indicate the presence of an interstellar civilization.
-In this scenario, artificial occulting objects would act as beacons, being placed in orbit around stars surrounding the central home star system.
-The orbital period of each artificial satellite would be proportional to the distance from the beacon star to the home star system.
-If the orbital period versus beacon distance relationship was known, the exact location of the home system could be determined via triangulation (or trilateration) from a subset of the transiting beacons seen by an observer.
-Current and future exoplanet surveys would be able to detect such spatially coordinated transits around nearby low-mass stars.
-While contrived, this example highlights the need for SETI researchers to consider the spatial-temporal domain.
-
 %\keywords{stars: activity --- stars: low-mass --- planets and satellites: detection}
@@
 Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.
-
-%\bf THE POINT: } in this new spatial-temporal domain, we should be looking for things that happen in space, such as over densities of events, or coordination of events like transits.
-
-%{\bf THIS PAPER NEEDS TO BE RETHOUGHT, IN LIGHT OF MORE SPECIFIC PAPER IDEA:}{\it SETI with ZTF}
-%
-%In the search for extraterrestrial intelligence (SETI), most approaches focus on direct detection of photons originating from transmission or waste energy.
-These searches occur over a range of wavelength regimes, and assume an extraterrestrial civilization will produce sufficient radiation to be detected apart from their parent star.
-While this may be feasible at radio wavelengths where Sun-like stars have lower emission, it is much more difficult to outshine the star in the optical regime where many present-day time domain surveys operate.
-%
-%
-%Blocking a significant fraction of light from a star may instead be much easier than outshining it \cite{arnold2005}.
-This has led to recent searches for transit-like signatures that may have artificial origins.
-For example, PAPER demonstrated the observable impact of transits by non-spherical objects \cite{arnold2005a}.
-%on the impact of artificial structures on transits. also this work on that:
-%\cite{wright2016}.
-%recent data from \Kepler \citep{borucki2010}, has found weird transit-like signals
-%\cite{boyajian2015}, 
-%which some have considered as SETI candidates. However, follow-up observations have yet to yield any confirming signatures, and instead this looks like a YSO w/ comets maybe \citep{lisse2015}.
-%
-%
-%Previous work has suggested looking for SETI signals from interstellar ``beacons''
-%\cite{benford2008}
-%However, most such work has been focused on using lasers or other means to out-shine a parent star, often observed using spectroscopy \cite{reines2002}.
-This is a very expensive way to stand out, and a slow way to find it, and so far has no success in discovering ET laser emission \cite{tellis2015}
-%\cite{arnold2005} note that objects could be placed at interesting spacing along the orbit to encode a pattern or simple message, demonstrating an intelligent origin.
-However, very little information can be effectively transmitted using eclipses, even with extreme precision in the recovery.
-%\cite{forgan2017} show a galactic-scale communications network of using transits, but note links in the network are only stable for order $10^5$ years as galactic orbital dynamics move things around
-
-%optical seti
-%\citep{holder2005} optical pulse SETI 
-%
-%a review of Schelling points from \citet{wright2017}
-%lighthouses \citep{zuluaga2015}
-%active and passive SETI in coordination (both temporal and spatial) with galactic events like supernovae 
-%\citep{lemarchand1994}
-%coordinated times/places, also known as Schelling or Focal points \citep{schelling1960}, are the key to finding people in unknown time/space.
-%planets in Earth's transit zone \cite{heller2016} - maybe transits would be aligned for our viewing benefit?
-%
-
-% include mantras of "worth doing", "do it almost for free", "reproducible", "portable, do it everywhere", and "outliers are nature papers"
 
@@
 \section{Example: Distribution of Transiting Exoplanets in the {\it Kepler} Field}
-%\section{Example 1: Coordinated Transits as ET Beacons}
 \label{sec:transit}
@@
 
-\citep{landy1993}
+The easiest conceptual method to search for a technosignature beacon in the spatial distribution of exoplanets would be looking for over-densities, i.e.
+regions with an unusually large number of transiting systems.
+However this requires a large and very uniform sampling of exoplanets per square degree to determine if the density is unusual.
+Such spatially uniform samples are difficult to achieve since e.g.
+bright stars can saturate large regions.
+
+Another simple statistical test is to measure the distribution of exoplanet systems using the two-point angular correlation function \citep{landy1993}.
+This method measures the probability of finding two sources separated by a given radial distance, and is often used in cosmology to characterize the distribution of galaxies on cosmological scales,.
+The two-point correlation function is easily measured thanks to generalized functions from packages such as {\tt astroML} \citep{astroMLText}, and  more efficient approaches have been produced that can scale to very large datasets \citep[e.g.][]{wang2013}.
+
+In Figure \ref{fig:2pt} I show the two-point correlation function for a sample of
 
@@
 
-%
-%\begin{figure}[!h]
-%\centering
-%\includegraphics[width=3in]{../figures/delta_per.png}
-%\caption{searching for correlation between transit periods and projected separations between all exoplanet systems out to 2deg in the Kepler data. 
-%median period differences are computed for bins of separation distance (orange points).
-%there is no trend in these medians, indicating a distribution of orbital periods that are random on large scales.}
-%\label{fig:2dcor}
-%\end{figure}
-
-%%%%%% 
-%    > shelve this idea for now, maybe another paper.... 1 demo is enough for this paper, 
-%    and it should be actually computed/worked out.
-%%%%%% 
-%\subsection{Coordinated Transit Networks}
-%imagine a Type N civilization with the ability to both travel to nearby star systems, and to build large enough structures in orbit around other stars to produce a visible transit in our data.
-%
-%give the detailed example i have thought up.
-%
-%becons placed at distances
-%
-%\begin{figure*}[]
-%\centering
-%\includegraphics[width=2.75in]{../figures/dist_per.pdf}
-%\includegraphics[width=2.75in]{../figures/sky_per.pdf}
-%\caption{schematic figure of the signal to detect in 2 dimensions. ra,dec in arbitrary units. red circle in middle is the home system}
-%\label{fig:2d}
-%\end{figure*}
-%
-%
-%for clarity, this is what the ideal signal might look like on the sky
-%
-%
-%this type of beacon network is advantageous because it points directly back towards their home. only a few systems actually need to be transiting from any given line of sight.
-%
-%NOW THE FULL SIMULATION...
-%- computation to do: if had 100 beacons, each placed at random orbital alignment, in 3d sphere around home system.
-%- assume G stars, 
-%- odds of observing transit of a fixed sized object versus orbital distance... goes down. need that plot to figure out probability. 
-%- assume ET places beacons with uniform RADIAL density in 3d space out to some maximum distance (even \# of systems as function of radial distance) in bins of 10pc out to 100 pc (i.e.
-10 beacons in each 10pc bin)
-%- orbital period is exact for each system, no bins of period
-%- do Monte Carlo sim with these parameters to figure out how many transiting systems we'd observe 
-%
-%---- make the plot for one MC realization of RA,Dec.... open circles for systems with no observed transits, colored for transits
-%
-%\begin{figure}[]
-%\centering
-%\includegraphics[width=3in]{../figures/3d_model.pdf}
-%\caption{the model in 3 dimensions projected into the sky plane.
-1000 simulated systems that span the galactic distance of 5-50 pc (blue) with 10 recovered transits highlighted (red).
-these would be hot Jupiters at the short P end.}
-%\label{fig:3d}
-%\end{figure}
-%
-%
-%IDEA:
-%could improve the efficiency of these beacons if we consider a bit of galactic structure, aligning the median orbital inclinations of the beacon systems with the Galactic plane, and only allowing a smaller range of possible inclinations.
-How much more efficient would it be if we forced $i<10^\circ$, instead of $i<90^\circ$?
-%
-%for a million system model w/ 90deg max, 0.46\% total detection efficiency. 
-%by constraining the inclination to 10deg max, get 4.3\% detection efficiency of planets. 30deg max gets 1.4\%
-%
-%
-%PROBLEM:
-%as \citet{forgan2017} note, an interstellar communications based on transits may only be stable for $\sim$10,000 years due to evolution of planetary orbits and relative motion between stars within the Milky Way.
-The beacon systems centered around the host system would have to be traveling as a moving group to maintain their respective trilateration signals, or have the transiting bodies actively over time.
-In other words, as the beacon systems moved relative to the ET host star due to differing Galactic orbits, the semi-major axis of the transiting bodies would have to be adjusted to compensate.
-%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The author thanks Nicholas M.
+J.R.A.D.
+thanks Nicholas M.
 Law and Jason T.
@@
 
-JRAD acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
+J.R.A.D.
+acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
 The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
+
+This paper includes data collected by the Kepler mission. Funding for the Kepler mission is provided by the NASA Science Mission directorate.
+
+This research has made use of the NASA Exoplanet Archive, which is operated by the California Institute of Technology, under contract with the National Aeronautics and Space Administration under the Exoplanet Exploration Program.
 

### 2019-04-25 paper/ms.tex
not much writing Thurs

@@
 
-In Figure \ref{fig:2pt} I show the two-point correlation function for a sample of
+In Figure \ref{fig:2pt} I show the two-point correlation function up to spatial scales of 5 degrees for both the sample of 2,324 transiting exoplanets discovered by the \Kepler primary mission,\footnote{Catalog from \url{http://exoplanetarchive.ipac.caltech.edu}} and for 201,312 stars in the \Kepler field.
+Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times.
+Here the correlation was done using only the (ra,dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g.
+including distance, orbital properties, etc).
 
-trivial example: two point correlation of exoplanets, Figure \ref{fig:2pt}
+%trivial example: two point correlation of exoplanets, Figure \ref{fig:2pt}
  not substantively different from field stars, so not a total engineering of the Kepler field.
@@
 
+This work made use of the gaia-kepler.fun crossmatch database created by Megan Bedell.
+
 \bibliography{/Users/james/Dropbox/references}

### 2019-05-01 paper/ms.tex
write thru section 5.1

@@
 The two examples outlined here explore the latter scenario, though I note the former is also well suited to automated approaches that would make use of future surveys and should be developed further.
+While I do not advocate these specific metrics are ideal for SETI, they demonstrate the type of signals we should design software systems to find in modern surveys.
 
@@
 
-%trivial example: two point correlation of exoplanets, Figure \ref{fig:2pt}
- not substantively different from field stars, so not a total engineering of the Kepler field.
-OK however note the selection function of these two populations is not exactly the same, so this can be a tricky metric to  use.
-considerable effort to ensure well understood sample and biases from non-uniform spatial selection, even in ``simple'' scenarios of galaxy populations in well-studied surveys \citep[e.g.][]{wang2013}.
-also requires large samples (like with simple spatial overdensities) to determine if something is unusual.
-This makes it an expensive beacon to create
+The correlation functions shown in Figure \ref{fig:2pt} are not substantively different between the exoplanet and field star populations, indicating no signs of a large-scale ``beacon'' based on the distribution of exoplanets in the \Kepler field.
+The slight differences in the correlation functions are not unexpected, given how sensitive this metric is to sample selection biases and non-uniformity.
+Indeed, even in ``simple'' scenarios of galaxy population analyses from well-studied surveys, considerable effort must be made to compare the correlation functions between different samples \citep[e.g.][]{wang2013}.
+As the 2-D correlation requires very homogeneous and large samples of sources to plausibly detect outliers with, it would be an inefficient metric for a civilization to base a beacon around.
 

### 2019-05-01 paper/ms.tex
start section 5.2

@@
 \subsection{Clusters of Similarity}
-more interesting example: how to stand out in exoplanet signals?
-if the simplest signal is just to indicate the presence of ET, then Q is how to stand out from the field?
-one idea is a spatial correlation of some exoplanet property, such as radius, eclipse depth, or orbital period.
-here i propose we search for spatial clusters of exoplanets with unusually similar orbital periods.
 
-to make an ideal transmitter of such a signal, the Q's are: what period to use, and how many systems need to be detected as similar to be confident something is odd about them?
+Rather than searching for an exoplanet beacon of outliers of systems in spatial density or distribution, we can consider clusters of systems that are unusually similar in some way.
+In the simplest case this could take the form of a cluster of transiting exoplanets with the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
+Here I demonstrate a simple algorithm for searching for clusters of exoplanets with unusually similar orbital periods.
+
+two questions arise: 
+- how many systems to use as a cluster, and how spatially tight should they be?
+- which periods and how self-similar do they need to be?
+
+%to make an ideal transmitter of such a signal, the Q's are: what period to use, and how many systems need to be detected as similar to be confident something is odd about them?
 the first Q a civilization might consider either periods that are unusual (long P i think, look at population histogram).
@@
 
-the second Q is akin to a loaded dice problem (i.e.
+%the second Q is akin to a loaded dice problem (i.e.
 how many rolls do you need to tell a die isnt rolling fair).
@@
 
-so for every planet in Kepler, I'll measure the.... stddev? sum of sq differences? some metric of "sameness"
-of the 4 (?) nearest exoplanets.
-then the Q is what's the  99\% outlier for the sameness distribution?
-choosing to do this simple numerical approach since the data are noisy, no way to exactly have ``same'' period....
-
-probably find no candidates, but that'll be AOK since the odds of 4 draws having exact same period should be very low.
+%so for every planet in Kepler, I'll measure the.... stddev? sum of sq differences? some metric of "sameness"
+%of the 4 (?) nearest exoplanets.
+%then the Q is what's the  99\% outlier for the sameness distribution?
+%choosing to do this simple numerical approach since the data are noisy, no way to exactly have ``same'' period....
+%probably find no candidates, but that'll be AOK since the odds of 4 draws having exact same period should be very low.
 

### 2019-05-09 paper/ms.tex
some words on the plane to tucson

@@
 Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation.
-Recent progress for systematic ``technosignature'' searchs has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.
+Recent progress for systematic ``technosignature'' searches has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.
 
@@
 While \citet{wright2018c} suggest the completeness of our search of the ``cosmic haystack'' at radio wavelengths is akin to the ratio of a swimming pool compared to the Earth's ocean's ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.
-However, new wide-field time domain surveys will enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing time resolved monitoring over many years.
+However, wide-field time domain surveys enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing time resolved monitoring over many years.
 
-Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI.
+\citet{djorgovski2000} highlighted the opportunity for a new age of generalized SETI using a virtual observatory made of digital sky surveys.
+New missions like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI.
 By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically.
@@
 In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
-The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon network.
+The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon.
 Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.
@@
 
-SETI with large surveys falls into two basic approaches: 1) classify every object or phenomena in the data, and any remaining outliers are technosignature candidates, or 2) search for specific ``impossible'' or unusual types of signals in the data, which I advocate for in this work.
+SETI with large surveys falls into two basic approaches: 1) classify every object or phenomena in the data, and any remaining outliers are technosignature candidates \citep{djorgovski2000}, or 2) search for specific ``impossible'' or unusual types of signals in the data, which I advocate for in this work.
 While the latter strategy does not guarantee a complete search for all SETI signals in the data, it is technically feasible for almost any time domain survey.
@@
 \subsection{Two-Point Correlation Function}
+\label{sec:2pt}
 
@@
 The two-point correlation function is easily measured thanks to generalized functions from packages such as {\tt astroML} \citep{astroMLText}, and  more efficient approaches have been produced that can scale to very large datasets \citep[e.g.][]{wang2013}.
-
-In Figure \ref{fig:2pt} I show the two-point correlation function up to spatial scales of 5 degrees for both the sample of 2,324 transiting exoplanets discovered by the \Kepler primary mission,\footnote{Catalog from \url{http://exoplanetarchive.ipac.caltech.edu}} and for 201,312 stars in the \Kepler field.
-Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times.
-Here the correlation was done using only the (ra,dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g.
-including distance, orbital properties, etc).
-
-The correlation functions shown in Figure \ref{fig:2pt} are not substantively different between the exoplanet and field star populations, indicating no signs of a large-scale ``beacon'' based on the distribution of exoplanets in the \Kepler field.
-The slight differences in the correlation functions are not unexpected, given how sensitive this metric is to sample selection biases and non-uniformity.
-Indeed, even in ``simple'' scenarios of galaxy population analyses from well-studied surveys, considerable effort must be made to compare the correlation functions between different samples \citep[e.g.][]{wang2013}.
-As the 2-D correlation requires very homogeneous and large samples of sources to plausibly detect outliers with, it would be an inefficient metric for a civilization to base a beacon around.
 
@@
 
+In Figure \ref{fig:2pt} I show the two-point correlation function up to spatial scales of 5 degrees for both the sample of 2,324 transiting exoplanets discovered by the \Kepler primary mission,\footnote{Catalog from \url{http://exoplanetarchive.ipac.caltech.edu}} and for 201,312 stars in the \Kepler field.
+Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times.
+Here the correlation was done using only the (ra,dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g.
+including distance, orbital properties, etc).
+
+The correlation functions shown in Figure \ref{fig:2pt} are not substantively different between the exoplanet and field star populations, indicating no signs of a large-scale ``beacon'' based on the distribution of exoplanets in the \Kepler field.
+The slight differences in the correlation functions are not unexpected, given how sensitive this metric is to sample selection biases and non-uniformity.
+Indeed, even in ``simple'' scenarios of galaxy population analyses from well-studied surveys, considerable effort must be made to compare the correlation functions between different samples \citep[e.g.][]{wang2013}.
+As the 2-D correlation requires very homogeneous and large samples of sources to plausibly detect outliers with, it would be an inefficient metric for a civilization to base a beacon around.
+
 %__________
@@
 
-Rather than searching for an exoplanet beacon of outliers of systems in spatial density or distribution, we can consider clusters of systems that are unusually similar in some way.
-In the simplest case this could take the form of a cluster of transiting exoplanets with the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
-Here I demonstrate a simple algorithm for searching for clusters of exoplanets with unusually similar orbital periods.
+Rather than searching for an exoplanet beacon consisting of outliers in spatial density or distribution (e.g.
+\S\ref{sec:2pt}), we can consider clusters of systems that are unusually similar in some way.
+In the simplest case this could take the form of a group of closely clustered stars with transiting exoplanets with the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
+The assumption here is that a civilization would have engineered this unusual set of transiting systems in order to be noticed by early astronomers (us) hoping to characterize the exoplanet demographics of our Galaxy.
+Here I demonstrate a simple algorithm for searching for such clusters of exoplanets with unusually similar orbital periods.
 
-two questions arise: 
-- how many systems to use as a cluster, and how spatially tight should they be?
-- which periods and how self-similar do they need to be?
+%two questions arise: 
+%- how many systems to use as a cluster, and how spatially tight should they be?
+%- which periods and how self-similar do they need to be?
+
+make clusters using easy KNN algorithm
 
@@
 \includegraphics[width=3in]{../figures/Kplanets.pdf}
-\caption{Spread in orbital periods for KNN clusters of exoplanet systems in \Kepler versus the average angular separation of the cluster.
-For $K=3$, 5 clusters have a spread in their periods below 6\% (red points).
-These KNN clusters have typical spatial separations of $\sim$0.1 deg.}
+\caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (ra, dec) space of transiting exoplanet systems in \Kepler.
+This sample only includes systems with only a single known transiting exoplanet.
+These KNN clusters have typical spatial separations of $\sim$0.1 deg.
+For $K$=3, five clusters have a spread in their orbital periods below 6\% (red points), but are not unusually close together in spatial separation.}
 \label{fig:knn}

### 2019-05-13 paper/ms.tex
writing on plane

@@
 \newcommand{\Kepler}{\textsl{Kepler}\xspace}
+\newcommand{\TESS}{\textsl{TESS}\xspace}
 
@@
 As a demonstration, I have computed the Cosmic Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}.
-This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}.
+This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[\TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}.
 These include space-based and traditional ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design.
@@
 To illustrate the potential for possible technosignature searches in the spatial-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets.
-Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or TESS, and they represent a natural ``Schelling point'' to attract attention with \citep{wright2017}.
+Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or \TESS, and they represent a natural ``Schelling point'' to attract attention with \citep{wright2017}.
 Creating such a beacon requires either generating a transiting system with unusual properties or patterns \citep[e.g.][]{arnold2005,kipping2016,forgan2017}, or as outlined in \S\ref{sec:strategy} an unusual {\it distribution} of transiting systems.
@@
 Rather than searching for an exoplanet beacon consisting of outliers in spatial density or distribution (e.g.
-\S\ref{sec:2pt}), we can consider clusters of systems that are unusually similar in some way.
-In the simplest case this could take the form of a group of closely clustered stars with transiting exoplanets with the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
-The assumption here is that a civilization would have engineered this unusual set of transiting systems in order to be noticed by early astronomers (us) hoping to characterize the exoplanet demographics of our Galaxy.
+\S\ref{sec:2pt}), we can consider groups of systems that are unusually similar in some way.
+In the simplest case this could take the form of many closely clustered stars with transiting exoplanets, where each planet had the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
+The assumption here is that a civilization would have engineered this unlikely set of transiting systems, like a lighthouse, in order to be noticed by early astronomers (us) hoping to characterize the exoplanet demographics of our Galaxy.
+This SETI approach relies on both the spatial and temporal information contained in exoplanet-hunting surveys like \Kepler or \TESS.
 Here I demonstrate a simple algorithm for searching for such clusters of exoplanets with unusually similar orbital periods.
@@
 
-make clusters using easy KNN algorithm
+As with many proposed technosignature signals, the key to this spatially distributed beacon system being effective is that the similarity metric for the cluster of stars must be substantially outside the realm of possible values that nature can produce.
+A civilization constructing such a beacon would also need to be aware of the natural distribution of exoplanet periods in order to determine the precision needed to place their transiting bodies.
+Thus we don't have to arbitrarily choose which orbital periods are interesting to search over or how close in period they need to be to count as a technosignature candidate, but instead we can adopt a data-driven approach where our thresholds are empirically set based on a probability threshold.
+
+For this demonstration I selected the same sample of 2,324 known transiting exoplanet systems within the \Kepler field used in \S\ref{sec:2pt}, with orbital periods ranging from 0.35 days to 1,322.5 days.
+I used the NearestNeighbor clustering algorithm in {\tt scikit-learn} for $k=3$ to easily identify and compute the projected angular distances between the nearest stars for each of the 2,324 exoplanet systems (i.e.
+an exoplanet and its two nearest neighbors).
+Even using the ``brute force'' algorithm, this clustering takes less than 1 second to compute for such small samples of objects.
+Since many of the exoplanets in the sample reside in multi-planet systems, the NearestNeighbor algorithm will produce clusters with multiple object separations of exactly zero.
+I therefore further restricted my sample to clusters that contain no multi-planet systems, leaving a total of 1,217 clusters constructed of three single transiting systems.
+Note also that due to the uniform but stochastic sampling of exoplanet systems across the field (see \S\ref{sec:2pt}), some clusters are partially or fully redundant (i.e.
+the same three stars can be selected in differing orders in up to three separate clusters).
+I do not resolve this redundancy in this demonstration.
+
+For each $k=3$ star cluster I compute two metrics that should be considered when searching  for these spatially distributed beacons, shown in Figure \ref{fig:knn}.
+First I simply calculate the mean cartesian distance of the three exoplanet  systems about the ``center'' or average spatial position.
+This enables us to potentially identify clusters of transiting systems that are unusually close together.
+Second I compute the similarity of the orbital periods for the three exoplanets in each clusters.
+This similarity metric here is defined as the range (i.e.
+max period -- min period) divided by the mean orbital period of the three stars.
+This period metric is easy to interpret, allowing us to identify clusters of transiting exoplanets whose orbital periods deviate by less than a given percentage.
 
@@
 These KNN clusters have typical spatial separations of $\sim$0.1 deg.
-For $K$=3, five clusters have a spread in their orbital periods below 6\% (red points), but are not unusually close together in spatial separation.}
+For $K$=3, three unique clusters have a spread in their orbital periods below 6\% (red points).
+However this is not a significant outlier in the orbital period similarity, and these clusters are not unusually close together in their spatial separation.}
 \label{fig:knn}
 \end{figure}
+
+We find no anonymously self-similar clusters of exoplanets in the \Kepler sample considered here.
+The cluster with the most similar period distribution consists of Kepler-1295, Kepler-1612, and Kepler-612.
+This triplet of systems have orbital periods within 5.1\% of each other (mean period of $P=3.82$ days).
 
@@
 %this kind of distributed network could be very efficient at broadcasting the presence of a civilization. dust clouds could create sufficient "transit" signals, while not affecting orbital dynamics
+
+while i've used self-similarity of clusters as the search metric here, we could instead pick densities of the most unlikely systems.
+Since long-period eclipses are geometrically the most rare, a spatial cluster of such exoplanets should immediately stand out as being very unlikely.
+However, these systems are also the most time-intensive to detect for an observer, and so I argue a more efficient beacon would be constructed of short-period eclipses.
+Adding the long-period or most unlikely cluster-finding algorithm to a future survey SETI toolkit is straightforward, however.
 
@@
 V.
-Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit.
+Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit, and the Research Corporation for Scientific Advancement for hosting the 2018 TDA Scialog meeting.
 

### 2019-05-13 paper/ms.tex
finish section5

@@
 The cluster with the most similar period distribution consists of Kepler-1295, Kepler-1612, and Kepler-612.
-This triplet of systems have orbital periods within 5.1\% of each other (mean period of $P=3.82$ days).
-
-%\citet{coughlin2014} find some correlations that are apparently removed in KOI catalog creation... due to contamination by RRlyr or other thing?
-%See also TCE histogram in \citet{twicken2016}
+This triplet of systems have orbital periods within 5.1\% of each other (mean period of $P=3.82$ days), only $2.4\sigma$ from the average of the period range distribution in Figure \ref{fig:knn}.
+This cluster also has a mean separation of 0.09 deg, only slightly closer than the typical cluster separation of 0.11 deg.
 

### 2019-05-16 paper/ms.tex
paper draft complete

@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{SETI in the Spatial-Temporal Survey Domain}
+\title{SETI in the Spatial-Temporal Survey Domain \\
+{\color{red} {\bf DRAFT} }
+}
 
@@
 Traditional searches for extraterrestrial intelligence (SETI) or ``technosignatures'' focus on dedicated observations of single stars or regions in the sky to detect excess or transient emission from intelligent sources.
-The latest generation of synoptic time domain surveys  enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
+The newest generation of synoptic time domain surveys enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
 Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space volume than many radio SETI investigations.
 Small-aperture, high cadence surveys like Evryscope can be comparable in their Cosmic Haystack volume completeness to deeper surveys including LSST.
-Investigations with these surveys can also be conducted at a fraction of the cost, since they require analysis of data already being gathered.
+Investigations with these surveys can also be conducted at a fraction of the cost of dedicated SETI surveys, since they require analysis of data already being gathered.
 However, SETI methodology has not widely utilized such surveys, and the literature is in need of new search algorithms that can incorporate signals from both the spatial and temporal domains.
-Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some new example SETI approaches that utilize the spatial-temporal domain
+Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some example SETI approaches using transiting exoplanets to form a distributed beacon.
 \end{abstract}
@@
 \subsection{Clusters of Similarity}
+\label{sec:sim}
 
@@
 
-For this demonstration I selected the same sample of 2,324 known transiting exoplanet systems within the \Kepler field used in \S\ref{sec:2pt}, with orbital periods ranging from 0.35 days to 1,322.5 days.
-I used the NearestNeighbor clustering algorithm in {\tt scikit-learn} for $k=3$ to easily identify and compute the projected angular distances between the nearest stars for each of the 2,324 exoplanet systems (i.e.
-an exoplanet and its two nearest neighbors).
+For this demonstration I selected the same sample used in \S\ref{sec:2pt} of 2,324 known transiting exoplanet systems within the \Kepler field, with orbital periods ranging from 0.35 days to 1,322.5 days.
+The NearestNeighbor clustering algorithm from {\tt scikit-learn} for $k=3$ was used to easily identify and compute the projected angular distances between the nearest stars for each of the 2,324 exoplanet systems (i.e.
+each exoplanet and its two nearest neighbors).
 Even using the ``brute force'' algorithm, this clustering takes less than 1 second to compute for such small samples of objects.
@@
 the same three stars can be selected in differing orders in up to three separate clusters).
-I do not resolve this redundancy in this demonstration.
+I do not eliminate this redundancy in this demonstration.
 
-For each $k=3$ star cluster I compute two metrics that should be considered when searching  for these spatially distributed beacons, shown in Figure \ref{fig:knn}.
+For each $k=3$ star cluster I compute two metrics that should be considered when searching for  spatially distributed technosignature beacons, shown in Figure \ref{fig:knn}.
 First I simply calculate the mean cartesian distance of the three exoplanet systems about the ``center'' or average spatial position.
-This enables us to potentially identify clusters of transiting systems that are unusually close together.
-Second I compute the similarity of the orbital periods for the three exoplanets in each clusters.
+This enables us to potentially identify clusters of transiting systems that are unusually close together in projected separation.
+Second I compute the similarity of the orbital periods for the three exoplanets in each cluster.
 This similarity metric here is defined as the range (i.e.
@@
 This period metric is easy to interpret, allowing us to identify clusters of transiting exoplanets whose orbital periods deviate by less than a given percentage.
-
-%to make an ideal transmitter of such a signal, the Q's are: what period to use, and how many systems need to be detected as similar to be confident something is odd about them?
-the first Q a civilization might consider either periods that are unusual (long P i think, look at population histogram).
-But to build such a beacon, might be efficient to balance with detectability since several systems must be detected to stand out, so short orbital period is probably best (easiest to detect).
-
-%the second Q is akin to a loaded dice problem (i.e.
-how many rolls do you need to tell a die isnt rolling fair).
-The odds of a single random transiting system being a given period is just a random draw from the underlying population, so the height of a bin in the normalized distribution.
-The odds of N draws is equivalent to multiplying the normalized distribution by itself N times.
-So looking at a new Figure i'm going to make, we need 4 (?) systems to be within the same bin to be highly unusual...
-use this to drive choice of N, and for a given result (if positive) estimate the probability
-
-%so for every planet in Kepler, I'll measure the.... stddev? sum of sq differences? some metric of "sameness"
-%of the 4 (?) nearest exoplanets.
-%then the Q is what's the  99\% outlier for the sameness distribution?
-%choosing to do this simple numerical approach since the data are noisy, no way to exactly have ``same'' period....
-%probably find no candidates, but that'll be AOK since the odds of 4 draws having exact same period should be very low.
 
@@
 
-We find no anonymously self-similar clusters of exoplanets in the \Kepler sample considered here.
+I find no anonymously self-similar $k=3$ clusters of exoplanets in the \Kepler sample considered here.
 The cluster with the most similar period distribution consists of Kepler-1295, Kepler-1612, and Kepler-612.
@@
 \label{sec:discussion}
-%
-%to get accurate census you need a complete time domain survey at some transit depth sensitivity out to some orbital period. going off our simulation, TESS might work for very short period planets
-%
-%TESS would be great for this in terms of spatial-temporal coverage. 
-%
-%LSST very good for finding SETI signals of this geometric style, but not ideal for events requiring such dedicating monitoring
-%
-%this kind of distributed network could be very efficient at broadcasting the presence of a civilization. dust clouds could create sufficient "transit" signals, while not affecting orbital dynamics
 
-while i've used self-similarity of clusters as the search metric here, we could instead pick densities of the most unlikely systems.
-Since long-period eclipses are geometrically the most rare, a spatial cluster of such exoplanets should immediately stand out as being very unlikely.
-However, these systems are also the most time-intensive to detect for an observer, and so I argue a more efficient beacon would be constructed of short-period eclipses.
-Adding the long-period or most unlikely cluster-finding algorithm to a future survey SETI toolkit is straightforward, however.
+In this paper I have presented the great potential for using algorithms to search for SETI or ``technosignature'' signals in modern spatial--temporal surveys, such as TESS, ZTF, or LSST.
+Though these surveys are primarily in optical wavelengths as opposed to traditional SETI at radio frequencies, there are many promising algorithmic approaches that may be considered, and these searches can ``piggyback'' on datasets and computational resources being developed for other science goals.
+This SETI approach also connects with the longterm rise of data science and statistical methods in astronomy to characterize, and find outliers in, large datasets.
+Interestingly, the search for technosignatures is possibly one of only a few areas where ``big data'' survey astronomy intersects with the growing astrobiology discipline.
+As \citet{berea2019} note, expertise in many complimentary domains of data analysis (e.g.
+cybersecurity) may also be useful for advancing survey SETI.
 
-so far we have focused on the need for developing SETI approaches that utilize the new spatial-temporal domain.
-The data streams from these surveys also enable many other search strategies that could be implemented.
-these might include:
-VASCO-type events, stars appearing or disappearing. \citep{villarroel2016}
-	- Esp. stars appearing that do not have a Gaia DR2 detection, for example.
-	- need to account for novae, etc...
-	- stars disappearing will be very interesting
+To illustrate the sorts of SETI algorithms that could be easily constructed based on existing data analysis and machine learning libraries, I have outlined a few simple approaches to search for beacons based on the distribution of exoplanets in the \Kepler field.
+The most detailed example in \S\ref{sec:sim} searches for clusters of transiting exoplanet systems with unusually similar properties.
+One obvious simplification of this approach as outlined is the need to pre-define the number of stars in each cluster, here assumed to be $k=3$.
+However, it is trivial in practice to adjust this number, or scan over many possible values.
+Indeed, I explored $k=2$ through $k=8$ clusters with this data, and found no substantial outliers in either the orbital period similarity or spatial separation.
 
-coordination/synchronization \citep{makovetskii1977,shostak2004}
-        - especially the SETI ellipsoid, a'la SN1987A \citep{lemarchand1994}
-        - could use galactic novae to provide time/place
-        - %https://asd.gsfc.nasa.gov/Koji.Mukai/novae/novae.html (new Novae)
-        - %http://www.cbat.eps.harvard.edu/nova_list.html historic Novae (till 2010)
+While I've explored self-similarity between spatially proximate exoplanet systems as the search metric here, one could instead pick densities of the {\it most unlikely} transiting systems.
+Since long-period eclipses are geometrically the most rare to detect, a spatial cluster of long-period exoplanets should immediately stand out as being very unlikely, even if the period similarity and average separation is not exceptional.
+Adding such a long-period or unlikely-period cluster-finding algorithm to a future survey SETI toolkit is straightforward.
+However, these systems are also the most time-intensive to detect for an observer, and so a more efficient beacon for both construction and detection would consist of short-period systems.
 
-unnaturals patterns of otherwise normal events
-	- look for e.g. flares or novae that recur w/ prime or fibonacci sequences
-	- see forthcoming paper
+As described in the introduction, these sky surveys enable a wide range of algorithmic approaches to use for SETI.
+Real-time alert streams from surveys like ZTF and soon LSST \citep[e.g.][]{schwamb2019} make rapid identification and follow-up of unusual variability straightforward to implement.
+This could include monitoring for over-densities in the spatial distribution of alerts, especially within the Earth Transit Zone \citep{heller2016}, for slowly appearing or disappearing stars on decades timescales \citep{villarroel2016}, or for coordinated variability with e.g.
+novae along the ``SETI ellipsoid'' over wide areas of the sky \citep{lemarchand1994}.
+In figure work we will explore a framework for searching the ZTF alert stream for unusual or ``impossible'' sequences of events.
 
-Surveys like ZTF (and soon LSST) publish real-time alerts of variability from sources. These alert streams could be monitored in real-time for signals as well:
-alerts from the stars within the ``restricted Earth Transit Zone'' \citep{heller2016}
-	- anytime we can see them w/ ZTF
-	- especially at our opposition (i.e. when we transit)
-	- with Gaia we can track stars that are leaving the ETZ as well, but might 
-	have seen us transit previously (or by the time the signal reaches us)
-
-alerts from known exoplanet host systems, especially coinciding with any known transit conjunction (mid-transit), 
-	- allow some tolerance window to account for TTVs
-
-The cost of deploying such SETI is remarkably low, since they can ``piggyback'' on databases and search tools being developed for the primary science drivers of each survey (e.g. transients for ZTF).
-
-%Lastly, we should resist the impulse to try and develop every possible search algorithm to mine this wealth of data for technosignatures.
-Instead, immediate progress can be made in SETI by deploying robust algorithms to search for specific signals
+Finally, while the opportunity for conducing SETI research using large surveys has been noted for nearly two decades \citep[e.g.][]{djorgovski2000}, what has been lacking is the development of general-use algorithms or software packages designed for such searches.
+This is an area in clear need of further exploration, and an opportunity for contribution to SETI by many software-minded researchers from all disciplines.
 

### 2019-05-29 paper/ms.tex
finished edits

@@
 %%%%%%%%%%%%%%%%%%%%%%
-\title{SETI in the Spatial-Temporal Survey Domain \\
+\title{SETI in the Spatio-Temporal Survey Domain \\
 {\color{red} {\bf DRAFT} }
@@
 Traditional searches for extraterrestrial intelligence (SETI) or ``technosignatures'' focus on dedicated observations of single stars or regions in the sky to detect excess or transient emission from intelligent sources.
-The newest generation of synoptic time domain surveys enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
+The newest generation of synoptic time domain surveys enable an entirely new approach: spatio--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time.
 Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space volume than many radio SETI investigations.
-Small-aperture, high cadence surveys like Evryscope can be comparable in their Cosmic Haystack volume completeness to deeper surveys including LSST.
-Investigations with these surveys can also be conducted at a fraction of the cost of dedicated SETI surveys, since they require analysis of data already being gathered.
-However, SETI methodology has not widely utilized such surveys, and the literature is in need of new search algorithms that can incorporate signals from both the spatial and temporal domains.
+Small-aperture, high cadence surveys like Evryscope can be comparable in their Haystack volume completeness to deeper surveys including LSST.
+Investigations with these surveys can also be conducted at a fraction of the cost of dedicated SETI surveys, since they make use of data already being gathered.
+However, SETI methodology has not widely utilized such surveys, and the field is in need of new search algorithms that can account for signals in both the spatial and temporal domains.
 Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some example SETI approaches using transiting exoplanets to form a distributed beacon.
@@
 
-Despite more than half a century of activity, the Search for Extra-Terrestrial Intelligence (SETI) is a field of study very much in its infancy. 
-\citet{wright2018c} note the volume of parameter space for SETI (or the search for a needle amongst the ``cosmic haystack'') at radio wavelengths has been barely explored, and that many obvious signals may indeed be awaiting our discovery.
-The lack of SETI activity by the astronomical community can be partially understood as due to funding limitations in recent decades, as well as social pressures against studying SETI experienced by professional astronomers and organizations \citep{wright2018b}.
+Despite more than half a century of activity, the Search for Extra-Terrestrial Intelligence (SETI) -- also described as the search for ``technosignatures'' -- is a field of study very much in its infancy.
+The lack of SETI research by the astronomical community is partially due to funding limitations in recent decades, as well as social pressures against studying SETI experienced by professional astronomers and organizations \citep{wright2018b}.
+Indeed, the volume of parameter space for SETI (or the search for a needle amongst the ``Cosmic Haystack'') at radio wavelengths has been barely explored \citet{wright2018c}, and thus many obvious signals may be awaiting our discovery.
+However, the new era of surveys and data-driven discovery in astronomy holds great promise for revolutionizing SETI.
 
-Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes.
+Traditional SETI work requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes.
 Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation.
@@
 
-The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When, where, and how to look?}''. 
-While \citet{wright2018c} suggest the completeness of our search of the ``cosmic haystack'' at radio wavelengths is akin to the ratio of a swimming pool compared to the Earth's ocean's ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.
-However, wide-field time domain surveys enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing time resolved monitoring over many years.
+The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When, where, and how}''. 
+While \citet{wright2018c} suggest the completeness of our search of the ``Cosmic Haystack'' at radio wavelengths is akin to the ratio of a small swimming pool compared to the Earth's oceans ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.
+Current optical-SETI is typically focused on searches for nanosecond pulses with specialized wide-field high-speed imaging \citep{howard2004}, or unusual spectroscopic features \citep{isaacson2019}.
+Wide-field time domain photometric surveys enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing precision time-resolved monitoring over many years.
 
 \citet{djorgovski2000} highlighted the opportunity for a new age of generalized SETI using a virtual observatory made of digital sky surveys.
-New missions like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI.
+New missions like the currently running Zwicky Transient Facility \citep[ZTF;][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST;][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI.
 By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically.
-As new search algorithms are developed, these databases can be quickly re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.
+As new search algorithms are developed, these databases can be re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.
 
@@
 In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches.
-The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon.
+The data-driven spatio-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon.
 Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.
@@
 By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look.
-Increasingly, multiple surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g.
-imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events..
-This also results in survey data providing some ``ground truth'' about the nature of most known objects in our Galaxy (e.g.
-their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} objects we're studying for SETI.
+Increasingly, multiple surveys obtain imaging that overlaps in both time and sky coverage, but with complementary properties (e.g.
+imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events.
+Surveys also produce important baseline information about the known objects in our Galaxy (e.g.
+their positions, temperatures, etc), which allows us to eliminate uncertainty on {\it what} objects being targeted for SETI.
 
@@
 Searches for new phenomena in time-domain astronomy can be carried out both in the future with new observations and in the past by re-analyzing survey archives.
-Using survey archives for SETI as opposed to targeted observations allows us to compare the results of competing algorithms or strategies, and enables reproducible science.
-Search algorithms are also portable, and ideally can be used to carry out SETI on any appropriately structured large survey.
+Using survey archives for SETI as opposed to targeted observations of individual stars allows us to compare the results of competing algorithms or strategies, and enables reproducible science.
+Search algorithms are also portable, and ideally can be used to carry out SETI on any new survey or large ensemble of observations.
 As future survey data is gathered, new classes of technosignature signals may be detectable that rely on decades worth of data. 
-Since any candidate technosignature detection will (or should) be the subject of great scrutiny, making the data and algorithms open and reproducible will be critical for comparison and independent validation.
+Since any technosignature detection candidate should be the subject of great scrutiny, making the data and algorithms open and reproducible will be critical for comparison and independent validation.
 
@@
 Large surveys, particularly ground-based efforts like LSST, are a highly prioritized component of the current and future development landscape in astronomy \citep{astro2010}.
-As such, great efforts are being made to develop survey tools and technologies to enable science from these missions, such as new database technologies \citep{juric2012}.
-These developments are especially needed to utilize the new forms of data being made available, including real-time data analysis \citep[e.g.][]{schwamb2019} and broadcasting real-time alerts for events \citep{patterson2019}.
+As such, great efforts are being made to develop survey tools and technologies to enable science from these missions, such as new database technologies \citep{juric2012,zecevic2019}, real-time event broadcasting \citep{patterson2019}, and real-time data analysis frameworks \citep[e.g.][]{schwamb2019}.
 The infrastructure needed to conduct technosignature searches with these large optical surveys is therefore already being developed.
-As with many areas of astrophysics in the ``survey era'', the biggest restriction to SETI with large surveys is therefore the availability of supported researchers mine the data.
+As with many areas of astrophysics in the ``survey era'', the biggest bottleneck for SETI with these surveys is the support (both financial and social) for researchers to develop algorithms and mine the available data.
 
@@
 
-Here I describe a few general classes of technosignature signals that may be well suited to developing new detection algorithms for spatial-temporal surveys:
+Here I describe a few general classes of technosignature signals that may be well suited to developing new detection algorithms for spatio-temporal surveys:
 
@@
 \item {\bf Spatial correlations of events or phenomena }-- e.g.
-coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse'' \citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
+coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse''  that defines the ellipsoid for receiving synchronized signals from multiple transmitters
+\citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
 %
-\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the Earth Transit Zone \citep{heller2016}, or in spatial clusters
+\item {\bf Spatial over-density or distribution} -- e.g.
+an over-density of a given phenomena, such as within the ``Earth Transit Zone'' band of stars that would see Earth as a transiting exoplanet \citep{heller2016}, or in spatial clusters
 %
-\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability, such as broadcasting a prime or fibonacci sequence of transits \citep{arnold2005,wright2016}.
+\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability (e.g.
+flares, pulsations, transits, etc), such as broadcasting a prime or fibonacci sequence using transits \citep{arnold2005,wright2016}.
 % 
 \end{enumerate}
+% omit piloted spacecraft, e.g. like looking for O'muamua objects
 
@@
 
+While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is important to quantitatively compare them to traditional searches in the radio.
 \citet{wright2018c} have provided a framework to quantify the volume of observable parameter space covered by various SETI projects -- dubbed the ``Cosmic Haystack''.
-The properties for several major radio SETI projects yield a Haystack volume coverage of $10^{-20}$ to $10^{-18}$, or as \citet{wright2018c} illustrate, akin to a drinking glass of water compared to the volume of Earth's oceans.
-While this tool was developed with radio surveys in mind, it is broadly applicable to surveys at other wavelengths that are not explicitly designed for SETI \citep[e.g.
+Though this metric was developed with radio surveys in mind, it is broadly applicable to surveys at other wavelengths that are not explicitly designed for SETI \citep[e.g.
 see][]{forgan2019}.
 
-While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is important to quantitatively compare them to traditional searches in the radio.
-For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Cosmic Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
+For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
 Since these terms all act as scalers in the Haystack volume integration, surveys with different properties (e.g.
@@
 For maximum impact, SETI should therefore be conducted on many types of surveys.
-%we need to be looking at big optical surveys, since they provide such amazing sky/time coverage
 
-As a demonstration, I have computed the Cosmic Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}.
+As a demonstration, I have computed the Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}.
 This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[\TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}.
-These include space-based and traditional ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design.
+These include space-based and ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design.
 I have assumed sensitivity ($S$) is the published per-visit photometric depth for each survey.
@@
 
-Typical values for the Haystack fraction for radio SETI programs as presented in \citet{wright2018c}  are in the range of $\log V_{Haystack} = $ 20 --17.
+Typical values for the Haystack fraction for radio SETI programs as presented in \citet{wright2018c}  are in the range of $\log V_{Haystack} =  -20$ to $-17$.
 However, these programs are mostly targeted observations of nearby stars.
@@
 
-By comparison, the wide field optical surveys I have featured all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction.
+The wide field optical surveys I have examined here all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction.
 Due to its very wide field coverage, and near continuous monitoring each night, Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) ranks as the best optical survey for technosignature searches of those surveyed here.
-The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Cosmic Haystack volume with as many different survey ``shapes'' as possible.
+The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Haystack volume with as many different survey ``shapes'' as possible.
+Note that since the area coverage, photometric depth, and cadence of these two surveys are so different, they will each be sensitive to very different specific SETI detection properties, such as signal repetition timescales or amplitudes.
 
-\citet{wright2018c} likened their Haystack fraction to searching for fish in a drinking glass of water compared to the Earth's oceans,.
+\citet{wright2018c} likened their Haystack fraction to searching for fish in a hot tub or small swimming pool compared to the Earth's oceans.
 Our results for SETI with optical surveys are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}.
@@
 
-To illustrate the potential for possible technosignature searches in the spatial-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets.
-Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or \TESS, and they represent a natural ``Schelling point'' to attract attention with \citep{wright2017}.
+To illustrate the potential for possible technosignature searches in the spatio-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets.
+Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or \TESS, and they represent a natural object to attract attention with, known as a ``Schelling point'' \citep{wright2017}.
 Creating such a beacon requires either generating a transiting system with unusual properties or patterns \citep[e.g.][]{arnold2005,kipping2016,forgan2017}, or as outlined in \S\ref{sec:strategy} an unusual {\it distribution} of transiting systems.
 The two examples outlined here explore the latter scenario, though I note the former is also well suited to automated approaches that would make use of future surveys and should be developed further.
-While I do not advocate these specific metrics are ideal for SETI, they demonstrate the type of signals we should design software systems to find in modern surveys.
+While I do not advocate these {\it specific} metrics are ideal for SETI, they demonstrate the type of signals we should design software systems to find in modern surveys.
 
@@
 The easiest conceptual method to search for a technosignature beacon in the spatial distribution of exoplanets would be looking for over-densities, i.e.
-regions with an unusually large number of transiting systems.
+regions with an unusually large number of transiting systems relative to the local density of stars.
 However this requires a large and very uniform sampling of exoplanets per square degree to determine if the density is unusual.
@@
 Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times.
-Here the correlation was done using only the (ra,dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g.
+Here the correlation was done using only the (RA, Dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g.
 including distance, orbital properties, etc).
@@
 In the simplest case this could take the form of many closely clustered stars with transiting exoplanets, where each planet had the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example.
-The assumption here is that a civilization would have engineered this unlikely set of transiting systems, like a lighthouse, in order to be noticed by early astronomers (us) hoping to characterize the exoplanet demographics of our Galaxy.
+The assumption here is that a civilization would have engineered this unlikely set of transiting systems, like a lighthouse, in order to be noticed by early astronomers who are characterizing the exoplanet demographics of our Galaxy (i.e.
+us).
 This SETI approach relies on both the spatial and temporal information contained in exoplanet-hunting surveys like \Kepler or \TESS.
 Here I demonstrate a simple algorithm for searching for such clusters of exoplanets with unusually similar orbital periods.
-
-%two questions arise: 
-%- how many systems to use as a cluster, and how spatially tight should they be?
-%- which periods and how self-similar do they need to be?
 
@@
 For each $k=3$ star cluster I compute two metrics that should be considered when searching for  spatially distributed technosignature beacons, shown in Figure \ref{fig:knn}.
-First I simply calculate the mean cartesian distance of the three exoplanet systems about the ``center'' or average spatial position.
+First I simply calculate the mean cartesian distance of the three exoplanet systems about the ``center'' or average spatial position in 2D space (RA, Dec).
 This enables us to potentially identify clusters of transiting systems that are unusually close together in projected separation.
@@
 \includegraphics[width=3in]{../figures/Kplanets.pdf}
-\caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (ra, dec) space of transiting exoplanet systems in \Kepler.
+\caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (RA, Dec) space of transiting exoplanet systems in \Kepler.
 This sample only includes systems with only a single known transiting exoplanet.
@@
 
-In this paper I have presented the great potential for using algorithms to search for SETI or ``technosignature'' signals in modern spatial--temporal surveys, such as TESS, ZTF, or LSST.
-Though these surveys are primarily in optical wavelengths as opposed to traditional SETI at radio frequencies, there are many promising algorithmic approaches that may be considered, and these searches can ``piggyback'' on datasets and computational resources being developed for other science goals.
-This SETI approach also connects with the longterm rise of data science and statistical methods in astronomy to characterize, and find outliers in, large datasets.
+In this paper I have presented the great potential for using algorithms to search for SETI or ``technosignature'' signals in modern spatio--temporal surveys, such as TESS, ZTF, or LSST.
+Though these surveys are primarily in optical wavelengths as opposed to traditional SETI at radio frequencies, there are many promising algorithmic approaches that may be considered.
+Further, these searches can ``piggyback'' on datasets and computational resources being developed for other science goals.
+This SETI approach also connects with the rise of data science and statistical methods in astronomy to characterize, and find outliers in, large datasets.
 Interestingly, the search for technosignatures is possibly one of only a few areas where ``big data'' survey astronomy intersects with the growing astrobiology discipline.
-As \citet{berea2019} note, expertise in many complimentary domains of data analysis (e.g.
+As \citet{berea2019} note, expertise in many complementary domains of data analysis (e.g.
 cybersecurity) may also be useful for advancing survey SETI.
@@
 
-As described in the introduction, these sky surveys enable a wide range of algorithmic approaches to use for SETI.
+As described in the introduction, sky surveys enable a wide range of algorithmic approaches to use for SETI.
 Real-time alert streams from surveys like ZTF and soon LSST \citep[e.g.][]{schwamb2019} make rapid identification and follow-up of unusual variability straightforward to implement.
-This could include monitoring for over-densities in the spatial distribution of alerts, especially within the Earth Transit Zone \citep{heller2016}, for slowly appearing or disappearing stars on decades timescales \citep{villarroel2016}, or for coordinated variability with e.g.
+Other approaches could include monitoring for over-densities in the spatial distribution of alerts, especially within the Earth Transit Zone \citep{heller2016}, searching for slowly appearing or disappearing stars on decades timescales \citep{villarroel2016}, or looking for coordinated variability with e.g.
 novae along the ``SETI ellipsoid'' over wide areas of the sky \citep{lemarchand1994}.
-In figure work we will explore a framework for searching the ZTF alert stream for unusual or ``impossible'' sequences of events.
+Unnatural motions of objects could also be searched for in sky surveys.
+This might be akin to searching for interstellar asteroids \citep[e.g.][]{mamajek2017} whose trajectories suddenly change.
+In future work we will explore a framework for searching the ZTF alert stream for unusual or ``impossible'' sequences of events in light curves.
 
@@
 This is an area in clear need of further exploration, and an opportunity for contribution to SETI by many software-minded researchers from all disciplines.
+
+\software{Python, IPython \citep{ipython}, NumPy \citep{numpy}, Matplotlib \citep{matplotlib}, SciPy \citep{scipy}, Pandas \citep{pandas}, Astropy \citep{astropy}}
 
@@
 Law and Jason T.
-Wright for their helpful ongoing conversations in developing this idea, and Andrew P.
+Wright for their ongoing conversations in developing this idea, Andrew P.
 V.
-Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit, and the Research Corporation for Scientific Advancement for hosting the 2018 TDA Scialog meeting.
+Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit, the Research Corporation for Scientific Advancement for hosting the 2018 TDA Scialog meeting, and Daniela Huppenkothen for comments that greatly improved this manuscript.
 

### 2019-05-29 paper/ms.tex
update figure names for submission
add bib

@@
 \end{abstract}
-
-%\keywords{stars: activity --- stars: low-mass --- planets and satellites: detection}
 
@@
 \centering
-\includegraphics[width=3.5in]{../figures/optical_haystack}
+\includegraphics[width=3.5in]{fig1}
 \caption{
@@
 \centering
-\includegraphics[width=3in]{../figures/2ptcf.pdf}
+\includegraphics[width=3in]{fig2}
 \caption{two point correlation function of transiting exoplanets versus all stars in the kepler field.
@@
 \centering
-\includegraphics[width=3in]{../figures/Kplanets.pdf}
+\includegraphics[width=3in]{fig3}
 \caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (RA, Dec) space of transiting exoplanet systems in \Kepler.
@@
 
-\bibliography{/Users/james/Dropbox/references}
+%\bibliography{/Users/james/Dropbox/references}
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Arnold}(2005)}]{arnold2005}
+{Arnold}, L. 2005, in SF2A-2005: Semaine de l'Astrophysique Francaise, ed.
+  F.~{Casoli}, T.~{Contini}, J.~M. {Hameury}, \& L.~{Pagani}, 207
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
+  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
+  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
+  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
+  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
+  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
+  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
+  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy}
+{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
+  \aap, 558, A33
+
+\bibitem[{{Bellm}(2014)}]{bellm2014}
+{Bellm}, E. 2014, in The Third Hot-wiring the Transient Universe Workshop, ed.
+  P.~R. {Wozniak}, M.~J. {Graham}, A.~A. {Mahabal}, \& R.~{Seaman}, 27--33
+
+\bibitem[{{Berea} {et~al.}(2019){Berea}, {Croft}, \& {Angerhausen}}]{berea2019}
+{Berea}, A., {Croft}, S., \& {Angerhausen}, D. 2019, arXiv e-prints,
+  arXiv:1903.08381
+
+\bibitem[{{Borucki} {et~al.}(2010){Borucki}, {Koch}, {Basri}, {Batalha},
+  {Brown}, {Caldwell}, {Caldwell}, {Christensen-Dalsgaard}, {Cochran},
+  {DeVore}, {Dunham}, {Dupree}, {Gautier}, {Geary}, {Gilliland}, {Gould},
+  {Howell}, {Jenkins}, {Kondo}, {Latham}, {Marcy}, {Meibom}, {Kjeldsen},
+  {Lissauer}, {Monet}, {Morrison}, {Sasselov}, {Tarter}, {Boss}, {Brownlee},
+  {Owen}, {Buzasi}, {Charbonneau}, {Doyle}, {Fortney}, {Ford}, {Holman},
+  {Seager}, {Steffen}, {Welsh}, {Rowe}, {Anderson}, {Buchhave}, {Ciardi},
+  {Walkowicz}, {Sherry}, {Horch}, {Isaacson}, {Everett}, {Fischer}, {Torres},
+  {Johnson}, {Endl}, {MacQueen}, {Bryson}, {Dotson}, {Haas}, {Kolodziejczak},
+  {Van Cleve}, {Chandrasekaran}, {Twicken}, {Quintana}, {Clarke}, {Allen},
+  {Li}, {Wu}, {Tenenbaum}, {Verner}, {Bruhweiler}, {Barnes}, \&
+  {Prsa}}]{borucki2010}
+{Borucki}, W.~J., {Koch}, D., {Basri}, G., {et~al.} 2010, Science, 327, 977
+
+\bibitem[{{Boyajian} {et~al.}(2015){Boyajian}, {LaCourse}, {Rappaport},
+  {Fabrycky}, {Fischer}, {Gandolfi}, {Kennedy}, {Liu}, {Moor}, {Olah}, {Vida},
+  {Wyatt}, {Best}, {Ciesla}, {Csak}, {Dupuy}, {Handler}, {Heng}, {Korhonen},
+  {Kovacs}, {Kozakis}, {Kriskovics}, {Schmitt}, {Szabo}, {Szabo}, {Wang},
+  {Goodman}, {Hoekstra}, \& {Jek}}]{boyajian2015}
+{Boyajian}, T.~S., {LaCourse}, D.~M., {Rappaport}, S.~A., {et~al.} 2015, ArXiv
+  e-prints, arXiv:1509.03622
+
+\bibitem[{Council(2010)}]{astro2010}
+Council, N.~R. 2010, New Worlds, New Horizons in Astronomy and Astrophysics
+  (Washington, DC: The National Academies Press), doi:10.17226/12951
+
+\bibitem[{{Djorgovski}(2000)}]{djorgovski2000}
+{Djorgovski}, S. 2000, in Astronomical Society of the Pacific Conference
+  Series, Vol. 213, Bioastronomy 99, ed. G.~{Lemarchand} \& K.~{Meech}, 519
+
+\bibitem[{{Djorgovski} {et~al.}(2013){Djorgovski}, {Mahabal}, {Drake},
+  {Graham}, \& {Donalek}}]{djorgovski2013}
+{Djorgovski}, S.~G., {Mahabal}, A., {Drake}, A., {Graham}, M., \& {Donalek}, C.
+  2013, {Sky Surveys}, ed. T.~D. {Oswalt} \& H.~E. {Bond}, 223
+
+\bibitem[{{Drake} {et~al.}(2009){Drake}, {Djorgovski}, {Mahabal}, {Beshore},
+  {Larson}, {Graham}, {Williams}, {Christensen}, {Catelan}, {Boattini},
+  {Gibbs}, {Hill}, \& {Kowalski}}]{catalina}
+{Drake}, A.~J., {Djorgovski}, S.~G., {Mahabal}, A., {et~al.} 2009, \apj, 696,
+  870
+
+\bibitem[{Forgan(2019)}]{forgan2019}
+Forgan, D. 2019, Research Notes of the {AAS}, 3, 32
+
+\bibitem[{{Forgan}(2017)}]{forgan2017}
+{Forgan}, D.~H. 2017, ArXiv e-prints, arXiv:1707.03730
+
+\bibitem[{{Heller} \& {Pudritz}(2016)}]{heller2016}
+{Heller}, R., \& {Pudritz}, R.~E. 2016, Astrobiology, 16, 259
+
+\bibitem[{{Howard} {et~al.}(2004){Howard}, {Horowitz}, {Wilkinson}, {Coldwell},
+  {Groth}, {Jarosik}, {Latham}, {Stefanik}, {Willman}, {Wolff}, \&
+  {Zajac}}]{howard2004}
+{Howard}, A.~W., {Horowitz}, P., {Wilkinson}, D.~T., {et~al.} 2004, \apj, 613,
+  1270
+
+\bibitem[{Hunter(2007)}]{matplotlib}
+Hunter, J.~D. 2007, Computing In Science \& Engineering, 9, 90
+
+\bibitem[{{Isaacson} {et~al.}(2019){Isaacson}, {Siemion}, {Marcy}, {Hickish},
+  {Price}, {Enriquez}, \& {Gizani}}]{isaacson2019}
+{Isaacson}, H., {Siemion}, A.~P.~V., {Marcy}, G.~W., {et~al.} 2019, \pasp, 131,
+  014201
+
+\bibitem[{{Isaacson} {et~al.}(2017){Isaacson}, {Siemion}, {Marcy}, {Lebofsky},
+  {Price}, {MacMahon}, {Croft}, {DeBoer}, {Hickish}, {Werthimer}, {Sheikh},
+  {Hellbourg}, \& {Enriquez}}]{isaacson2017}
+---. 2017, \pasp, 129, 054501
+
+\bibitem[{{Ivezi{\'c}} {et~al.}(2014){Ivezi{\'c}}, {Connolly}, {Vanderplas}, \&
+  {Gray}}]{astroMLText}
+{Ivezi{\'c}}, {\v Z}., {Connolly}, A., {Vanderplas}, J., \& {Gray}, A. 2014,
+  Statistics, Data Mining and Machine Learning in Astronomy (Princeton
+  University Press)
+
+\bibitem[{{Ivezi{\'c}} {et~al.}(2008){Ivezi{\'c}}, {Tyson}, {Acosta},
+  {Allsman}, {Anderson}, {Andrew}, {Angel}, {Axelrod}, {Barr}, {Becker},
+  {Becla}, {Beldica}, {Blandford}, {Bloom}, {Borne}, {Brandt}, {Brown},
+  {Bullock}, {Burke}, {Chandrasekharan}, {Chesley}, {Claver}, {Connolly},
+  {Cook}, {Cooray}, {Covey}, {Cribbs}, {Cutri}, {Daues}, {Delgado}, {Ferguson},
+  {Gawiser}, {Geary}, {Gee}, {Geha}, {Gibson}, {Gilmore}, {Gressler}, {Hogan},
+  {Huffer}, {Jacoby}, {Jain}, {Jernigan}, {Jones}, {Juric}, {Kahn}, {Kalirai},
+  {Kantor}, {Kessler}, {Kirkby}, {Knox}, {Krabbendam}, {Krughoff}, {Kulkarni},
+  {Lambert}, {Levine}, {Liang}, {Lim}, {Lupton}, {Marshall}, {Marshall}, {May},
+  {Miller}, {Mills}, {Monet}, {Neill}, {Nordby}, {O'Connor}, {Oliver},
+  {Olivier}, {Olsen}, {Owen}, {Peterson}, {Petry}, {Pierfederici},
+  {Pietrowicz}, {Pike}, {Pinto}, {Plante}, {Radeka}, {Rasmussen}, {Ridgway},
+  {Rosing}, {Saha}, {Schalk}, {Schindler}, {Schneider}, {Schumacher}, {Sebag},
+  {Seppala}, {Shipsey}, {Silvestri}, {Smith}, {Smith}, {Strauss}, {Stubbs},
+  {Sweeney}, {Szalay}, {Thaler}, {Vanden Berk}, {Walkowicz}, {Warner},
+  {Willman}, {Wittman}, {Wolff}, {Wood-Vasey}, {Yoachim}, {Zhan}, \& {for the
+  LSST Collaboration}}]{lsst}
+{Ivezi{\'c}}, {\v Z}., {Tyson}, J.~A., {Acosta}, E., {et~al.} 2008, ArXiv
+  e-prints, \# 0805.2366, arXiv:0805.2366
+
+\bibitem[{Jones {et~al.}(2001--)Jones, Oliphant, Peterson, {et~al.}}]{scipy}
+Jones, E., Oliphant, T., Peterson, P., {et~al.} 2001--, {SciPy}: Open source
+  scientific tools for {Python}, , , [Online; accessed <today>]
+
+\bibitem[{{Juric}(2012)}]{juric2012}
+{Juric}, M. 2012, {LSD: Large Survey Database framework}, , , ascl:1209.003
+
+\bibitem[{{Kipping} \& {Teachey}(2016)}]{kipping2016}
+{Kipping}, D.~M., \& {Teachey}, A. 2016, \mnras, 459, 1233
+
+\bibitem[{{Lacki}(2019)}]{lacki2019}
+{Lacki}, B.~C. 2019, arXiv e-prints, arXiv:1903.05839
+
+\bibitem[{{Landy} \& {Szalay}(1993)}]{landy1993}
+{Landy}, S.~D., \& {Szalay}, A.~S. 1993, \apj, 412, 64
+
+\bibitem[{{Law} {et~al.}(2015){Law}, {Fors}, {Ratzloff}, {Wulfken},
+  {Kavanaugh}, {Sitar}, {Pruett}, {Birchard}, {Barlow}, {Cannon}, {Cenko},
+  {Dunlap}, {Kraus}, \& {Maccarone}}]{law2015}
+{Law}, N.~M., {Fors}, O., {Ratzloff}, J., {et~al.} 2015, \pasp, 127, 234
+
+\bibitem[{{Lemarchand}(1994)}]{lemarchand1994}
+{Lemarchand}, G.~A. 1994, \apss, 214, 209
+
+\bibitem[{{Makovetskii}(1977)}]{makovetskii1977}
+{Makovetskii}, P.~V. 1977, \sovast, 21, 251
+
+\bibitem[{{Mamajek}(2017)}]{mamajek2017}
+{Mamajek}, E. 2017, Research Notes of the American Astronomical Society, 1, 21
+
+\bibitem[{McKinney(2010)}]{pandas}
+McKinney, W. 2010, in Proceedings of the 9th Python in Science Conference, ed.
+  S.~van~der Walt \& J.~Millman, 51 -- 56
+
+\bibitem[{Oliphant(2007)}]{numpy}
+Oliphant, T.~E. 2007, Computing in Science Engineering, 9, 10
+
+\bibitem[{{Patterson} {et~al.}(2019){Patterson}, {Bellm}, {Rusholme}, {Masci},
+  {Juric}, {Krughoff}, {Golkhou}, {Graham}, {Kulkarni}, {Helou}, \& {Zwicky
+  Transient Facility Collaboration}}]{patterson2019}
+{Patterson}, M.~T., {Bellm}, E.~C., {Rusholme}, B., {et~al.} 2019, \pasp, 131,
+  018001
+
+\bibitem[{P\'erez \& Granger(2007)}]{ipython}
+P\'erez, F., \& Granger, B.~E. 2007, Computing in Science and Engineering, 9,
+  21
+
+\bibitem[{{Price} {et~al.}(2018){Price}, {MacMahon}, {Lebofsky}, {Croft},
+  {DeBoer}, {Enriquez}, {Foster}, {Gajjar}, {Hellbourg}, {Isaacson}, {Siemion},
+  {Werthimer}, {Green}, {Amy}, {Ball}, {Bock}, {Craig}, {Edwards}, {Jameson},
+  {Mader}, {Preisig}, {Smith}, {Reynolds}, \& {Sarkissian}}]{price2018}
+{Price}, D.~C., {MacMahon}, D.~H.~E., {Lebofsky}, M., {et~al.} 2018, ArXiv
+  e-prints, arXiv:1804.04571
+
+\bibitem[{{Ricker} {et~al.}(2015){Ricker}, {Winn}, {Vanderspek}, {Latham},
+  {Bakos}, {Bean}, {Berta-Thompson}, {Brown}, {Buchhave}, {Butler}, {Butler},
+  {Chaplin}, {Charbonneau}, {Christensen-Dalsgaard}, {Clampin}, {Deming},
+  {Doty}, {De Lee}, {Dressing}, {Dunham}, {Endl}, {Fressin}, {Ge}, {Henning},
+  {Holman}, {Howard}, {Ida}, {Jenkins}, {Jernigan}, {Johnson}, {Kaltenegger},
+  {Kawai}, {Kjeldsen}, {Laughlin}, {Levine}, {Lin}, {Lissauer}, {MacQueen},
+  {Marcy}, {McCullough}, {Morton}, {Narita}, {Paegert}, {Palle}, {Pepe},
+  {Pepper}, {Quirrenbach}, {Rinehart}, {Sasselov}, {Sato}, {Seager},
+  {Sozzetti}, {Stassun}, {Sullivan}, {Szentgyorgyi}, {Torres}, {Udry}, \&
+  {Villasenor}}]{tess}
+{Ricker}, G.~R., {Winn}, J.~N., {Vanderspek}, R., {et~al.} 2015, Journal of
+  Astronomical Telescopes, Instruments, and Systems, 1, 014003
+
+\bibitem[{Schwamb {et~al.}(2019)Schwamb, Hsieh, Bannister, Bodewits, Chesley,
+  Fraser, Granvik, Jones, Juri{\'{c}}, Kelley, Ragozzine, Trilling, \&
+  and}]{schwamb2019}
+Schwamb, M.~E., Hsieh, H., Bannister, M.~T., {et~al.} 2019, Research Notes of
+  the {AAS}, 3, 51
+
+\bibitem[{{Shostak}(2004)}]{shostak2004}
+{Shostak}, S. 2004, in IAU Symposium, Vol. 213, Bioastronomy 2002: Life Among
+  the Stars, ed. R.~{Norris} \& F.~{Stootman}, 409
+
+\bibitem[{{Tarter}(2001)}]{tarter2001}
+{Tarter}, J. 2001, \araa, 39, 511
+
+\bibitem[{{Tingay} {et~al.}(2018){Tingay}, {Tremblay}, \& {Croft}}]{tingay2018}
+{Tingay}, S.~J., {Tremblay}, C.~D., \& {Croft}, S. 2018, \apj, 856, 31
+
+\bibitem[{{Villarroel} {et~al.}(2016){Villarroel}, {Imaz}, \&
+  {Bergstedt}}]{villarroel2016}
+{Villarroel}, B., {Imaz}, I., \& {Bergstedt}, J. 2016, \aj, 152, 76
+
+\bibitem[{{Wang} {et~al.}(2013){Wang}, {Brunner}, \& {Dolence}}]{wang2013}
+{Wang}, Y., {Brunner}, R.~J., \& {Dolence}, J.~C. 2013, \mnras, 432, 1961
+
+\bibitem[{{Worden} {et~al.}(2017){Worden}, {Drew}, {Siemion}, {Werthimer},
+  {DeBoer}, {Croft}, {MacMahon}, {Lebofsky}, {Isaacson}, {Hickish}, {Price},
+  {Gajjar}, \& {Wright}}]{worden2017}
+{Worden}, S.~P., {Drew}, J., {Siemion}, A., {et~al.} 2017, Acta Astronautica,
+  139, 98
+
+\bibitem[{{Wright}(2017)}]{wright2017}
+{Wright}, J.~T. 2017, {Exoplanets and SETI}, 186
+
+\bibitem[{{Wright} {et~al.}(2016){Wright}, {Cartier}, {Zhao}, {Jontof-Hutter},
+  \& {Ford}}]{wright2016}
+{Wright}, J.~T., {Cartier}, K.~M.~S., {Zhao}, M., {Jontof-Hutter}, D., \&
+  {Ford}, E.~B. 2016, \apj, 816, 17
+
+\bibitem[{{Wright} {et~al.}(2018){Wright}, {Kanodia}, \& {Lubar}}]{wright2018c}
+{Wright}, J.~T., {Kanodia}, S., \& {Lubar}, E. 2018, \aj, 156, 260
+
+\bibitem[{{Wright} \& {Oman-Reagan}(2018)}]{wright2018b}
+{Wright}, J.~T., \& {Oman-Reagan}, M.~P. 2018, International Journal of
+  Astrobiology, 17, 177
+
+\bibitem[{{Ze{\v{c}}evi{\'c}} {et~al.}(2019){Ze{\v{c}}evi{\'c}}, {Slater},
+  {Juri{\'c}}, {Connolly}, {Lon{\v{c}}ari{\'c}}, {Bellm}, {Golkhou}, \&
+  {Suberlak}}]{zecevic2019}
+{Ze{\v{c}}evi{\'c}}, P., {Slater}, C.~T., {Juri{\'c}}, M., {et~al.} 2019, arXiv
+  e-prints, arXiv:1905.09034
+
+\end{thebibliography}
 
